---
title: "Galapagos_map"
author: "Cheryl Logan"
date: "9/10/2020"
output: html_document
---

Make a map of Galapagos SSTs using data from https://www.bio-oracle.org/code.php

The data available in Bio-ORACLE are documented in two peer reviewed articles that you should cite:

Tyberghein L, Verbruggen H, Pauly K, Troupin C, Mineur F, De Clerck O (2012) Bio-ORACLE: A global environmental dataset for marine species distribution modelling. Global Ecology and Biogeography, 21, 272–281.
[access publication]   [supporting information]

Assis, J., Tyberghein, L., Bosh, S., Verbruggen, H., Serrão, E. A., & De Clerck, O. (2017). Bio-ORACLE v2.0: Extending marine data layers for bioclimatic modelling. Global Ecology and Biogeography.
[access publication]   [supporting information]


## Install packages
```{r setup, include=FALSE}
#install.packages("sdmpredictors")
#install.packages("leaflet")
```

## Load Libraries
```{r setup, include=FALSE}
library(sdmpredictors)
library(leaflet)
```

## Load Raster data

```{r sdm datasets}
# Explore datasets in the package
list_datasets()
list_layers()
```

## Load SST data

```{r sdm datasets}
sst <- load_layers(c("BO_sstmax", "BO_sstmean", "BO_sstmin", "BO_sstrange"))

#Spatial location and grain
#Global at 5 arcmin (c. 0.08° or 9.2 km at the equator).

#Time period and grain
#Present (2000–2014) environmental conditions based on monthly averages.
```
##  Plots

```{r glps sst, echo=FALSE}
# Easy download of raster file (sst mean)
sst.mean <- load_layers("BO_sstmean")
sst.max <- load_layers("BO_sstmax")
sst.min <- load_layers("BO_sstmin")
sst.range <- load_layers("BO_sstrange")

# Crop raster to fit the Galapagos is
glps.ext <- extent(-94,-88, -3,3)
sst.mean.crop <- crop(sst.mean, glps.ext)
sst.max.crop <- crop(sst.max, glps.ext)
sst.min.crop <- crop(sst.min, glps.ext)
sst.range.crop <- crop(sst.range, glps.ext)

# Generate a nice color ramp and plot the map
plot(sst.mean.crop,col=colorRampPalette(c("blue","purple", "yellow","red"))(255), axes=FALSE, box=FALSE)
title(cex.sub = 1.25, sub = "Mean SST (ºC)")

```
# Sampling Sites 
```{r}
#Sites

#Darwin
points(cbind(-92,1.67), pch=21, cex=1, bg="red") 
text(x=-92.7,y=1.7, labels="Darwin")

#Wolf
points(cbind(-91.8,1.38), pch=21, cex=1, bg="red") 
text(x=-92.3,y=1.2, labels="Wolf")

#Isabela
points(cbind(-90.9,-.9), pch=21, cex=1, bg="red") 
text(x=-91.5,y=-1.2, labels="Isabela")

#Floreana
points(cbind(-90.47,-1.25), pch=21, cex=1, bg="red") 
text(x=-90.8,y=-1.7, labels="Floreana")

#Espanola
points(cbind(-89.68,-1.3), pch=21, cex=1, bg="red") 
text(x=-89.1,y=-1.7, labels="Espanola")

#San Cristobal
points(cbind(-89.4,-.72), pch=21, cex=1, bg="red") 
text(x=-89,y=-.5, labels="San Cristobal")

```

##  Plot Galapagos Is

```{r sst, echo=FALSE}

#pdf(file="sst.mean.crop.pdf") # uncomment to print to pdf

plot(sst.mean.crop,main="Mean SST - Galapagos Islands", col=colorRampPalette(c("blue","purple", "yellow","red"))(255), 
legend.args=list(text='Degrees C', side=4,font=2, line=2.5, cex=0.8),axes=TRUE, box=FALSE)

#San Cristobal
points(cbind(-89.4,-.72), pch=21, cex=2, bg="red") 
text(x=-89,y=-.5, labels="San Cristobal")

#Darwin
points(cbind(-92,1.67), pch=21, cex=2, bg="red") 
text(x=-92.7,y=1.7, labels="Darwin")

#Wolf
points(cbind(-91.8,1.38), pch=21, cex=2, bg="red") 
text(x=-92.3,y=1.2, labels="Wolf")

#Isabela
points(cbind(-90.9,-.9), pch=21, cex=2, bg="red") 
text(x=-91.5,y=-1.2, labels="Isabela")

#Espanola
points(cbind(-89.68,-1.3), pch=21, cex=2, bg="red") 
text(x=-89.1,y=-1.7, labels="Espanola")

#Floreana
points(cbind(-90.47,-1.25), pch=21, cex=2, bg="red") 
text(x=-90.8,y=-1.7, labels="Floreana")

#dev.off() # uncomment to print to pdf

```
##  Plot SST Range - Galapagos Is

```{r sst, echo=FALSE}

#pdf(file="sst.range.crop.pdf")  # uncomment to print to pdf

plot(sst.range.crop,main="SST Range - Galapagos Islands", col=colorRampPalette(c("blue","purple", "yellow","red"))(255), 
legend.args=list(text='Delta degC', side=4,font=2, line=2.5, cex=0.8),axes=TRUE, box=TRUE)

#San Cristobal
points(cbind(-89.4,-.72), pch=21, cex=2, bg="red") 
text(x=-89,y=-.5, labels="San Cristobal")

#Darwin
points(cbind(-92,1.67), pch=21, cex=2, bg="red") 
text(x=-92.7,y=1.7, labels="Darwin")

#Wolf
points(cbind(-91.8,1.38), pch=21, cex=2, bg="red") 
text(x=-92.3,y=1.2, labels="Wolf")

#Isabela
points(cbind(-90.9,-.9), pch=21, cex=2, bg="red") 
text(x=-91.5,y=-1.2, labels="Isabela")

#Espanola
points(cbind(-89.68,-1.3), pch=21, cex=2, bg="red") 
text(x=-89.1,y=-1.7, labels="Espanola")

#Floreana
points(cbind(-90.43,-1.3), pch=21, cex=2, bg="red") 
text(x=-90.8,y=-1.7, labels="Floreana")

#dev.off() # uncomment to print to pdf
```


##  Plot Min SST Galapagos Is

```{r sst, echo=FALSE}
plot(sst.min.crop,main="Min SST - Galapagos Islands", col=colorRampPalette(c("blue","purple", "yellow","red"))(255), 
legend.args=list(text='Degrees C', side=4,font=2, line=2.5, cex=0.8),axes=FALSE, box=FALSE)

#San Cristobal
points(cbind(-89.4,-.72), pch=21, cex=1, bg="red") 
text(x=-89,y=-.5, labels="San Cristobal")

#Darwin
points(cbind(-92,1.67), pch=21, cex=1, bg="red") 
text(x=-92.7,y=1.7, labels="Darwin")

#Wolf
points(cbind(-91.8,1.38), pch=21, cex=1, bg="red") 
text(x=-92.3,y=1.2, labels="Wolf")

#Isabela
points(cbind(-90.9,-.9), pch=21, cex=1, bg="red") 
text(x=-91.5,y=-1.2, labels="Isabela")

#Espanola
points(cbind(-89.68,-1.3), pch=21, cex=1, bg="red") 
text(x=-89.1,y=-1.7, labels="Espanola")

#Floreana
points(cbind(-90.43,-1.3), pch=21, cex=1, bg="red") 
text(x=-90.8,y=-1.7, labels="Floreana")
```

##  Plot Max SST Galapagos Is

```{r sst, echo=FALSE}
plot(sst.max.crop,main="Max SST - Galapagos Islands", col=colorRampPalette(c("blue","purple", "yellow","red"))(255), 
legend.args=list(text='Degrees C', side=4,font=2, line=2.5, cex=0.8),axes=FALSE, box=FALSE)

#Darwin
points(cbind(-92,1.67), pch=21, cex=1, bg="red") 
text(x=-92.7,y=1.7, labels="Darwin")

#Wolf
points(cbind(-91.8,1.38), pch=21, cex=1, bg="red") 
text(x=-92.3,y=1.2, labels="Wolf")

#Isabela
points(cbind(-90.9,-.9), pch=21, cex=1, bg="red") 
text(x=-91.5,y=-1.2, labels="Isabela")

#Floreana
points(cbind(-90.43,-1.3), pch=21, cex=1, bg="red") 
text(x=-90.8,y=-1.7, labels="Floreana")

#Espanola
points(cbind(-89.68,-1.3), pch=21, cex=1, bg="red") 
text(x=-89.1,y=-1.7, labels="Espanola")

#San Cristobal
points(cbind(-89.4,-.72), pch=21, cex=1, bg="red") 
text(x=-89,y=-.5, labels="San Cristobal")

```


```{r sst, echo=FALSE}
#Sample code 2 : Extracting environmental information for a set of sites
# Load packages (leaflet allows to load google maps)
library(sdmpredictors)
library(leaflet)

# List layers avaialble in Bio-ORACLE v2
layers.bio2 <- list_layers( datasets="Bio-ORACLE" )
layers.bio2

# Download environmental data layers (Mean SST, Max SST, Min SST, SST range)
SST.glps <- load_layers( layercodes = c("BO_sstmean" , "BO_sstmax", "BO_sstmin", "BO_sstrange") , equalarea=FALSE, rasterstack=TRUE)

# Generate a data.frame with the sites of interest
my.sites <- data.frame(Name=c("Darwin","Wolf","Isabela","Floreana","Espanola","San Cristobal") , Lon=c(-92.7,-92.3,-91.5,-90.8,-89.1,-89) , Lat=c(1.7,1.2,-1.2,-1.3,-1.3,-.72) )
my.sites

# Visualise sites of interest in google maps
m <- leaflet()
m <- addTiles(m)
m <- addMarkers(m, lng=my.sites$Lon, lat=my.sites$Lat, popup=my.sites$Name)
m

# Extract environmental values from layers
my.sites.environment <- data.frame(Name=my.sites$Name) 
extract(SST.glps,my.sites[,2:3]) 
my.sites.environment
```