---
title: "Galapagos_PAM_allsites"
author: "Cheryl Logan"
date: "4/1/2019"
output: html_document
---

## Load libraries
```{r setup, include = FALSE}
library(tidyverse)
library(ggplot2)
library(RColorBrewer)
library(ggpubr)
library(thermPerf)
library(nlme)
```

## Load data (from "Galapagos_PAM_allsites.Rmd")
```{r load data, include = FALSE}
PAM_all <- read_csv("/Users/loga8761/github/PAM/Galapagos_all.csv")
PAM_cold <- read_csv("/Users/loga8761/github/PAM/Galapagos_cold.csv")
PAM_heat <- read_csv("/Users/loga8761/github/PAM/Galapagos_heat.csv")
```


## Compare CoralWatch Color Change with PAM Change
```{r color v PAM}
# using ggpubr to make the plot
sp <- ggscatter(PAM_all, x = "ColorChange", y = "PAMChange",
   color = "Species", 
   add = "reg.line", conf.int = TRUE)
sp + stat_cor(aes(color = Species), label.x = -4) +
   ggtitle("PAM Change versus Color Change - All Sites") +
  theme_bw() +
  ylab("Change in Photochemical Efficency (Fv/Fm)") +
  xlab("CoralWatch Color Change") +
  ggsave("PAM_Color_allsites_scatter.png", width = 7, height = 5)

# using ggplot2 to make the plot
# ggplot(PAM_all, aes(x = ColorChange, y = PAM)) +
#   geom_point() +
#   aes(color = Species) +
#   facet_wrap(~ Species) +
#   theme_bw() +
#   ylab("Photochemical Efficency (Fv/Fm)") +
#   xlab("CoralWatch Color Change") +
#   stat_cor(method = "pearson", label.x = 2.5, label.y = .2)
#   ggtitle("PAM versus CoralWatch Color Change - All Sites") +
#   ggsave("PAM_Color_allsites_scatter.png", width = 7, height = 5)
#   #stat_smooth(method = "lm", formula = y ~ x) + # try linear model
#   
#   ggtitle("PAM versus CoralWatch Color Change - All Sites") +
#   ggsave("PAM_Color_allsites_scatter.png", width = , height = 5)

summary(lm(PAM ~ ColorChange, PAM_all))
```

## Compare Color Change by Temperature
```{r color density}
# remove data from Floreana 31C temp & Wolf 26C from one tank each
ggplot(PAM_all, aes(x = ColorChange)) +
  geom_density() +
  aes(color = Species) +
  facet_wrap(~ Treatment) +
  theme_bw() +
  ylab("Frequency") +
  xlab("Color Change") +
  xlim(c(-4,0)) +
  ggtitle("Temperature versus CoralWatch Color Change - All Sites") +
  ggsave("ColorvTempHistogram_allsites_scatter.png", width = 7, height = 5)
```

## Scatterplot with lm fit by species
```{r jitterplot}
# add second axis to plot Color and PAM on same figure
ggplot(PAM_all, aes(x = Treatment, y = PAMChange)) +
  geom_point(alpha = 0.3, color = "#009E73",colour = "PAM") +
  geom_smooth(method = "loess", mapping = aes(color = RampType)) +
  
  #geom_smooth(method = "lm", formula = y ~ exp(x), fill = "gray", colour = "black") + # try linear model
  #geom_smooth(method = "lm", formula = y ~ log(x), fill = "green", colour = "blue") + # try linear model
  facet_wrap(~Species) +
  theme_bw() +
  facet_grid(Species ~ Site) +
  ylab("PAM (Fv/Fm) (green)") +
  xlab("Temperature [°C]") +
  ggtitle("PAM - All Sites") +
  ggsave("PAM_allsites_scatter_mod.png", width = 7, height = 5)

#test
# ggplot(PAM_all, aes(x = Treatment, y = PAMChange)) + geom_point() +
#   geom_point(data=B,colour='red') + xlim(0, 10) 
```


## Boxplots by species & site 
```{r boxplot}

# PAM Change plot
ggplot(PAM_all,aes(x = Treatment, y = PAMChange, color = Species)) +
  geom_boxplot(outlier.size = 0.3) +
  aes(group = Treatment) +
  facet_grid(Species ~ Site) +
  ylab("Change in Photochemical Efficency (Fv/Fm)") +
  xlab("Temperature [°C]") +
  theme_bw() +
  ggtitle("PAM Fluorometry - All Sites") +
  ggsave("PAM_allsites_boxplot.png", width = 7, height = 3)

# Color Change plot
ggplot(PAM_all,aes(x = Treatment, y = ColorChange, color = Species)) +
  geom_boxplot(outlier.size = 0.3) +
  aes(group = Treatment) +
  facet_grid(Species ~ Site) +
  ylab("Visual Color Change (pre vs post stress)") +
  xlab("Temperature [°C]") +
  theme_bw() +
  ggtitle("Visual Color Change - All Sites") +
  ggsave("Color_allsites_boxplot.png", width = 7, height = 3)
```
## Boxplots by species & site 
```{r boxplot}
# PAM Change plot
ggplot(PAM_all,aes(x = Treatment, y = PAMChange, color = Site)) +
  geom_jitter(outlier.size = 0.3) +
  #aes(group = Treatment) +
  #facet_grid(Species ~ Site) +
  ylab("Change in Photochemical Efficency (Fv/Fm)") +
  xlab("Temperature [°C]") +
  theme_bw() 
  #ggtitle("PAM Fluorometry - All Sites") +
  #ggsave("PAM_allsites_boxplot.png", width = 7, height = 3)

# Color Change plot
ggplot(PAM_all,aes(x = Treatment, y = ColorChange, color = Species)) +
  geom_point(outlier.size = 0.3) +
  facet_grid(Species ~ Site) +
  ylab("Visual Color Change (pre vs post stress)") +
  xlab("Temperature [°C]") +
  theme_bw() 
  #ggtitle("Visual Color Change - All Sites") +
  #ggsave("Color_allsites_boxplot.png", width = 7, height = 3)
```

## Plots by ramp type
```{r PAM plot by ramp type}
# Heat Ramp
ggplot(PAM_heat,aes(x = Treatment, y = PAMChange, color = Species)) +
  geom_jitter() +
  facet_grid(Species ~ Site) +
  stat_smooth(method = "lm", formula = y ~ exp(x), fill = "gray", colour = "black") + # try linear model
  ylab("Photochemical Efficency (Fv/Fm)") +
  xlab("Treatment Temperature (degC)") +
  theme_bw() +
  theme(legend.position = "none") +
  ggtitle("Heat Ramp PAM All Sites") +
  ggsave("PAM_heatonly_allsites_fit.png", width = 7, height = 3)

# Cold Ramp
ggplot(PAM_cold,aes(x = Treatment, y = PAMChange, color = Species)) +
  geom_jitter() +
  facet_grid(Species ~ Site) +
  stat_smooth(method = "lm", formula = y ~ log(x), fill = "gray", colour = "black") + # try linear model
  ylab("Photochemical Efficency (Fv/Fm)") +
  xlab("Treatment Temperature (degC)") +
  theme_bw() +
  theme(legend.position = "none") +
  ggtitle("Cold Ramp PAM All Sites") +
  ggsave("PAM_coldonly_allsites_fit.png", width = 7, height = 3)
```

## Plots by ramp type for simposio talk
```{r PAM plot simposio}

PAM_heat$Site = factor(PAM_heat$Site, levels=c('Darwin', 'Wolf2', 'Isabela','Punta_Pitt', 'Espanola'))
PAM_cold$Site = factor(PAM_cold$Site, levels=c('Darwin', 'Wolf2', 'Isabela','Punta_Pitt', 'Espanola'))

PAM_heat <- PAM_heat %>% filter(Site %in% c("Darwin", "Wolf2", "Isabela", "Punta_Pitt", "Espanola"))

PAM_cold <- PAM_cold %>%
  filter(Site %in% c("Darwin", "Wolf2", "Isabela", "Punta_Pitt", "Espanola"))
 

ggplot(PAM_heat,aes(x = Treatment, y = PAMChange, color = Species)) +
  geom_jitter() +
  facet_grid(Species ~ Site) +
  stat_smooth(method = "lm", formula = y ~ exp(x), fill = "gray", colour = "black") + # try linear model
  ylab("Photochemical Efficency (Fv/Fm)") +
  xlab("Treatment Temperature (degC)") +
  ylim(c(-.4,0)) +
  theme_bw() +
  theme(legend.position = "none") +
  ggtitle("Heat Ramp PAM All Sites") +
  ggsave("PAM_heatonly_allsites_fit.png", width = 7, height = 3)

ggplot(PAM_cold,aes(x = Treatment, y = PAMChange, color = Species)) +
  geom_jitter() +
  facet_grid(Species ~ Site) +
  stat_smooth(method = "lm", formula = y ~ log(x), fill = "gray", colour = "black") + # try linear model
  ylab("Photochemical Efficency (Fv/Fm)") +
  xlab("Treatment Temperature (degC)") +
  ylim(c(-.4,.1)) +
  theme_bw() +
  theme(legend.position = "none") +
  ggtitle("Cold Ramp PAM All Sites") +
  ggsave("PAM_coldonly_allsites_fit.png", width = 7, height = 3)
```


```{r PAM plot by ramp type}
ggplot(PAM_heat,aes(x = Treatment, y = ColorChange, color = Species)) +
  geom_jitter() +
  facet_grid(Species ~ Site) +
  stat_smooth(method = "lm", formula = y ~ exp(x), colour = "black",fill = "black") + # try linear model
  ylab("CoralWatch Color Change") +
  xlab("Treatment Temperature (degC)") +
  theme_bw() +
  theme(legend.position = "none") +
  ggtitle("Heat Ramp Color All Sites") +
  ggsave("Color_heatonly_allsites_fit.png", width = 7, height = 3)

ggplot(PAM_cold,aes(x = Treatment, y = ColorChange, color = Species)) +
  geom_jitter() +
  facet_grid(Species ~ Site) +
  stat_smooth(method = "lm", formula = y ~ log(x), colour = "black", fill = "black") + # try linear model
  ylab("CoralWatch Color Change") +
  xlab("Treatment Temperature (degC)") +
  theme_bw() +
  theme(legend.position = "none") +
  ggtitle("Cold Ramp Color All Sites") +
  ggsave("Color_coldonly_allsites_fit.png", width = 7, height = 3)
```

## PAM-Exponential model fitting & stats
```{r lm models}
mod_heat <- lm(PAMChange ~ exp(Treatment) * Species * Site, data = PAM_heat)
summary(mod_heat)

mod_cold <- lm(PAMChange ~ log(Treatment) * Species * Site, data = PAM_cold)
summary(mod_cold)
```

## Color-Exponential model fitting & stats
```{r lm models}
mod_heat <- lm(ColorChange ~ exp(Treatment) * Species * Site, data = PAM_heat)
summary(mod_heat)

mod_cold <- lm(ColorChange ~ log(Treatment) * Species * Site, data = PAM_cold)
summary(mod_cold)
```

## PAM Pocillopora ANOVA model fitting & Tukey HSD
```{r ANOVA stats}

PAM_all_poc <- PAM_all %>%
  filter(Species == "Pocillopora_sp") 

excludes <- PAM_all %>%
  filter(Treatment == 14 |Treatment == 31)

PAM_all_poc <- anti_join(PAM_all_poc, excludes)


#Pocillopora
mod_anova_all <- aov(PAMChange ~ factor(Site), data = PAM_all_poc)
summary(mod_anova_all)
TukeyHSD(mod_anova_all)

ggplot(PAM_all_poc,aes(x = Site, y = PAMChange)) +
  geom_boxplot(outlier.size = 0.3) +
  facet_wrap(. ~ Treatment) +
  ylab("Normalized Change in Photochemical Efficency (Fv/Fm)") +
  xlab("Pocillopora by Site") +
  theme_bw() +
  #ylim(c(.4,.6)) +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust=1)) +
  ggsave("Poc_by_site.png", width = 7, height = 6)
```


## PAM Pavona ANOVA model fitting & Tukey HSD
```{r ANOVA stats}
PAM_all_pav <- PAM_all %>%
  filter(Species == "Pavona_clavus") 

excludes <- PAM_all %>%
  filter(Treatment == 14 |Treatment == 31)

PAM_all_pav <- anti_join(PAM_all_pav, excludes)


#Pavona
mod_anova_all <- aov(PAMChange ~ factor(Treatment) * Site, data = PAM_all_pav)
summary(mod_anova_all)
TukeyHSD(mod_anova_all)

ggplot(PAM_all_pav,aes(x = Site, y = PAMChange)) +
  geom_boxplot(outlier.size = 0.3) +
  facet_wrap(. ~ Treatment) +
  ylab("Normalized Change in Photochemical Efficency (Fv/Fm)") +
  xlab("Pavona by Site") +
  theme_bw() +
  #ylim(c(.4,.6)) +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust=1)) +
  ggsave("Pav_by_site.png", width = 7, height = 6)

##HSD_PAM_aov <- TukeyHSD(mod_anova_all)
#names(HSD_PAM_aov)
#test <- (HSD_PAM_aov$`factor(Treatment):Species:Site`)

#write_csv(as.data.frame(test), "test")

```


## Color N way ANOVA model fitting & Tukey HSD
```{r ANOVA stats}
mod_anova_all <- aov(ColorChange ~ factor(Treatment) * Species * Site, data = PAM_all)
summary(mod_anova_all)
HSD_Color <- TukeyHSD(mod_anova_all)
```
