---
title: "Galapagos_PAM_allsites"
author: "Cheryl Logan"
date: "4/1/2019"
output: html_document
---

## Load libraries
```{r setup, include = FALSE}
library(tidyverse)
library(ggplot2)
library(RColorBrewer)
library(ggpubr)
library(thermPerf)
```

## Load data
```{r load data, include = FALSE}
wd <- "/Users/loga8761/Google Drive/Galapagos/CBASS_Galapagos/PAM/PAM_allsites_forR/PAM_allsites_forR_color/"

Flo_heat <- read_csv(file=paste0(wd,"2019_03_17_heat_Floreana_foranalysis.csv"))
# remove data from Floreana 31C temp (messed up temp profile for that run)
excludes <- Flo_heat %>%
  filter(Treatment == 31, Replicate == 'A')
Flo_heat <- anti_join(Flo_heat, excludes)

Flo_cold <- read_csv(file=paste0(wd,"2019_03_18_cold_Floreana_foranalysis.csv"))

Isa_heat <- read_csv(file=paste0(wd,"2019_03_19_heat_Isabela_foranalysis.csv"))
Isa_cold <- read_csv(file=paste0(wd,"2019_03_20_cold_Isabela_foranalysis.csv"))

Pitt_heat <- read_csv(file=paste0(wd,"2019_03_26_heat_PuntaPitt_foranalysis.csv"))
Pitt_cold <- read_csv(file=paste0(wd,"2019_03_27_cold_PuntaPitt_foranalysis.csv"))

Esp_heat <- read_csv(file=paste0(wd,"2019_04_02_heat_Espanola_foranalysis.csv"))
Esp_cold <- read_csv(file=paste0(wd,"2019_04_03_cold_Espanola_foranalysis.csv"))
```

## Join files into new data frame
```{r join data, include = FALSE}
PAM_all <- bind_rows(Flo_heat, Flo_cold, Isa_heat, Isa_cold, Pitt_heat, Pitt_cold, Esp_heat, Esp_cold)
PAM_heat <- bind_rows(Flo_heat, Isa_heat, Pitt_heat, Esp_heat)
PAM_cold <- bind_rows(Flo_cold, Isa_cold, Pitt_cold, Esp_cold)

PAM_all <- PAM_all %>%
  select(Site, Species, Treatment, PAM, Replicate, Geno, Color_1, Color_2) %>%
  mutate(Site = as.factor(Site)) %>%
  mutate(ColorChange = Color_2 - Color_1)

PAM_heat <- PAM_heat %>%
  select(Site, Species, Treatment, PAM, Replicate, Geno, Color_1, Color_2) %>%
  mutate(Site = as.factor(Site)) %>%
  mutate(ColorChange = Color_2 - Color_1)

PAM_cold <- PAM_cold %>%
  select(Site, Species, Treatment, PAM, Replicate, Geno, Color_1, Color_2) %>%
  mutate(Site = as.factor(Site)) %>%
  mutate(ColorChange = Color_2 - Color_1)
```

## Compare CoralWatch Color Change with PAM
```{r color v PAM}
# using ggpubr to make the plot
sp <- ggscatter(PAM_all, x = "ColorChange", y = "PAM",
   color = "Species", 
   add = "reg.line", conf.int = TRUE)
sp + stat_cor(aes(color = Species), label.x = -4) +
   ggtitle("PAM versus CoralWatch Color Change - All Sites") +
  ylim(0, .8) +
  theme_bw() +
  ylab("Photochemical Efficency (Fv/Fm)") +
  xlab("CoralWatch Color Change") +
  ggsave("PAM_Color_allsites_scatter.png", width = 7, height = 5)

# using ggplot2 to make the plot
# ggplot(PAM_all, aes(x = ColorChange, y = PAM)) +
#   geom_point() +
#   aes(color = Species) +
#   facet_wrap(~ Species) +
#   theme_bw() +
#   ylab("Photochemical Efficency (Fv/Fm)") +
#   xlab("CoralWatch Color Change") +
#   stat_cor(method = "pearson", label.x = 2.5, label.y = .2)
#   ggtitle("PAM versus CoralWatch Color Change - All Sites") +
#   ggsave("PAM_Color_allsites_scatter.png", width = 7, height = 5)
#   #stat_smooth(method = "lm", formula = y ~ x) + # try linear model
#   
#   ggtitle("PAM versus CoralWatch Color Change - All Sites") +
#   ggsave("PAM_Color_allsites_scatter.png", width = , height = 5)

summary(lm(PAM ~ ColorChange, PAM_all))
```

## Compare CoralWatch Color Change with PAM
```{r color density}
# remove data from Floreana 31C temp
ggplot(PAM_all, aes(x = ColorChange)) +
  geom_density() +
  aes(color = Species) +
  facet_wrap(~ Treatment) +
  theme_bw() +
  ylab("Frequency") +
  xlab("Color Change") +
  xlim(c(-4,0)) +
  ggtitle("Temperature versus CoralWatch Color Change - All Sites") +
  ggsave("ColorvTempHistogram_allsites_scatter.png", width = 7, height = 5)
```

## Scatterplot with lm fit by species
```{r jitterplot}
# add second axis to plot Color and PAM on same figure
ggplot(PAM_all, aes(x = Treatment, y = PAM)) +
  geom_jitter(alpha = 0.3, color = "#009E73",colour = "PAM") +
  theme(axis.title.y = element_text(colour = "#009E73")) +
  

  #add color data
  geom_jitter(aes(y = ((ColorChange + 1)/6)),alpha = 0.3, color = "steelblue") +
  scale_y_continuous(sec.axis = sec_axis(~.*6 - 1, name = "CoralWatch Color Change (blue)")) +
  theme(axis.title.y = element_text(colour = "steelblue")) +
  
  theme_bw() +
  facet_grid(Species ~ Site) +
  ylab("PAM (Fv/Fm) (green)") +
  xlab("Temperature [°C]") +
  
  ggtitle("PAM and Color All Sites") +
  ggsave("PAMvColor_allsites_scatter.png", width = 7, height = 5)
```


## Boxplot by species & site
```{r boxplot}
ggplot(PAM_all,aes(x = Treatment, y = PAM, color = Species)) +
  geom_boxplot(outlier.size = 0.3) +
  aes(group = Treatment) +
  facet_grid(Species ~ Site) +
  ylab("Photochemical Efficency (Fv/Fm)") +
  xlab("Treatment Temperature (degC)") +
  theme_bw() +
  ggtitle("PAM All Sites") +
  ggsave("PAM_allsites_boxplot.png", width = 7, height = 3)
```

## Plots by ramp type
```{r PAM plot by ramp type}
ggplot(PAM_heat,aes(x = Treatment, y = PAM, color = Species)) +
  geom_jitter() +
  facet_grid(Species ~ Site) +
  stat_smooth(method = "lm", formula = y ~ exp(x)) + # try linear model
  ylab("Photochemical Efficency (Fv/Fm)") +
  xlab("Treatment Temperature (degC)") +
  theme_bw() +
  theme(legend.position = "none") +
  ggtitle("Heat Ramp PAM All Sites") +
  ggsave("PAM_heatonly_allsites_fit.png", width = 7, height = 3)

ggplot(PAM_cold,aes(x = Treatment, y = PAM, color = Species)) +
  geom_jitter() +
  facet_grid(Species ~ Site) +
  stat_smooth(method = "lm", formula = y ~ log(x)) + # try linear model
  ylab("Photochemical Efficency (Fv/Fm)") +
  xlab("Treatment Temperature (degC)") +
  theme_bw() +
  theme(legend.position = "none") +
  ggtitle("Cold Ramp PAM All Sites") +
  ggsave("PAM_coldonly_allsites_fit.png", width = 7, height = 3)
```
```{r PAM plot by ramp type}
ggplot(PAM_heat,aes(x = Treatment, y = ColorChange, color = Species)) +
  geom_jitter() +
  facet_grid(Species ~ Site) +
  stat_smooth(method = "lm", formula = y ~ exp(x)) + # try linear model
  ylab("CoralWatch Color Change") +
  xlab("Treatment Temperature (degC)") +
  theme_bw() +
  theme(legend.position = "none") +
  ggtitle("Heat Ramp Color All Sites") +
  ggsave("Color_heatonly_allsites_fit.png", width = 7, height = 3)

ggplot(PAM_cold,aes(x = Treatment, y = ColorChange, color = Species)) +
  geom_jitter() +
  facet_grid(Species ~ Site) +
  stat_smooth(method = "lm", formula = y ~ log(x)) + # try linear model
  ylab("CoralWatch Color Change") +
  xlab("Treatment Temperature (degC)") +
  theme_bw() +
  theme(legend.position = "none") +
  ggtitle("Cold Ramp Color All Sites") +
  ggsave("Color_coldonly_allsites_fit.png", width = 7, height = 3)
```

## PAM-Exponential model fitting & stats
```{r lm models}
mod_heat <- lm(PAM ~ exp(Treatment) * Species * Site, data = PAM_heat)
summary(mod_heat)

mod_cold <- lm(PAM ~ log(Treatment) * Species * Site, data = PAM_cold)
summary(mod_cold)
```

## Color-Exponential model fitting & stats
```{r lm models}
mod_heat <- lm(ColorChange ~ exp(Treatment) * Species * Site, data = PAM_heat)
summary(mod_heat)

mod_cold <- lm(ColorChange ~ log(Treatment) * Species * Site, data = PAM_cold)
summary(mod_cold)
```

## PAM 2 way ANOVA model fitting & Tukey HSD
```{r ANOVA stats}
mod_anova_all <- aov(PAM ~ factor(Treatment) * Species * Site, data = PAM_all)
summary(mod_anova_all)
TukeyHSD(mod_anova_all)
```


## Color 2 way ANOVA model fitting & Tukey HSD
```{r ANOVA stats}
mod_anova_all <- aov(ColorChange ~ factor(Treatment) * Species * Site, data = PAM_all)
summary(mod_anova_all)
TukeyHSD(mod_anova_all)
```
