---
title: "GLPS_stats_forCOAST"
author: "Cheryl Logan"
date: "9/11/2020"
output: html_document
---

## Load libraries
```{r setup, include = FALSE}
#library(thermPerf)
#install.packages("ggrepel")
library(tidyverse)
library(ggplot2)
library(RColorBrewer)
library(ggpubr)
library(ggrepel)
library(nlme)
library(tidyverse)
library(ggpubr)
library(rstatix)
library(forcats)
library(ggthemes)
library(lme4)
library(lmerTest)
library(emmeans)
library(car)
```

## Load data (from "Galapagos_PAM_allsites.Rmd")
```{r load data, include = FALSE}
PAM_all <- read_csv("/Users/loga8761/github/PAM/Galapagos_all.csv")
PAM_cold <- read_csv("/Users/loga8761/github/PAM/Galapagos_cold.csv")
PAM_heat <- read_csv("/Users/loga8761/github/PAM/Galapagos_heat.csv")
```

## Rename sites and remove bad data
```{r rename sites, include = FALSE}
# Change site names to islands and remove sites with bad data
pam.data <- PAM_all %>%                             
  mutate_at("Site",str_replace,"Wolf2", "Wolf") %>%             
  mutate_at("Site",str_replace,"AcademyBay", "Santa Cruz") %>%  
  mutate_at("Site",str_replace,"Punta_Pitt", "San Cristobal") %>%
  mutate(Site = fct_relevel(Site, 
            "Darwin", "Wolf", "San Cristobal", 
            "Espanola", "Isabela"))  %>%
  filter(Site != "Santa Cruz"& Site != "Floreana") %>% # Remove sites with data issues
  filter(Treatment != "14"& Site != "31") # Remove 14C and 31C temps since only used at Floreana

PAM_all_poc <- pam.data %>% # select Pocillopora only
  filter(Species == "Pocillopora_sp") 
```

# ANOVA stats - old, from 2020 COAST proposal
```{r COAST stats, include=FALSE}
#Load the data and inspect one random row by group combinations:
PAM_all_poc %>% sample_n_by(Treatment, Site, Replicate, size = 1)

#Compute the mean and the standard deviation (SD) of PAM by groups:
PAM_all_poc %>%
  group_by(Treatment, Site, Replicate) %>%
  get_summary_stats(PAM, type = "mean_sd")

#Create a box plot of PAM by treatment
#note Isabela rep needs to be removed
bxp <- ggboxplot(
  PAM_all_poc, x = "Treatment", y = "PAM", 
  color = "Replicate", palette = "jco", facet.by = "Site"
  )
bxp

#ANOVA
res.aov <- PAM_all_poc %>% anova_test(PAM ~ Treatment*Site*Replicate)
res.aov

#Pocillopora
mod_anova_all <- aov(PAMChange ~ factor(Treatment)*Site*Replicate, data = PAM_all_poc)
summary(mod_anova_all)
TukeyHSD(mod_anova_all)

```

## Violin Plot used for 2020 COAST Proposal (do not change!)
```{r,echo=TRUE}
# Labels for re-ordering site names
SiteLabels <- c("Darwin", "Wolf","San Cristobal", "Espanola", "Isabela")

# Create violin plot for both species across all sites
ggplot(pam.data,aes(x = Site, y = PAMChange, fill = Site) )+
  geom_violin(scale="width",outlier.size = 0.3, alpha=0.8) +
  geom_boxplot(width=0.1, color="black", alpha=0.2) +
  facet_grid(Species ~ Treatment) +
  ylab("Change in FvFm") +
  theme_bw() +
  scale_x_discrete(labels= SiteLabels) +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust=1)) 

# Save with date stamp
ggsave(paste0("/Users/loga8761/github/PAM/Figures/PAM_Violin_by_site_"
              , format(Sys.time(), "%Y-%m-%d_%H-%M")
              , ".png"), width = 7, height = 4) # save as .png
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

Create a plot with fraction of most heat- and cold- tolerant corals by site
(Note: re-did this analysis using lm in Colony Tolerance Rmd)
```{r sd most tolerant, echo=FALSE}
# For Pocillopora
PAM_all_poc_clean <- PAM_all_poc %>%
  select(PAM, PAMChange,ColorChange, Color_2,Site, Species,Treatment,Geno,Replicate)
  
# percent difference = abs((Value1 - Value2)/Value1)*100
df <- group_by(PAM_all_poc_clean,Site, Species,Treatment,Geno) %>%
  arrange(Site,Treatment,desc(Geno)) %>% # sort so replicates are next to each other
  mutate(geno_mean = mean(PAMChange, na.rm = TRUE)) %>% # genotype mean
  mutate(pct_diff = abs((PAMChange[1] - PAMChange[2])/PAMChange[1])*100, na.rm = TRUE)  
    
# Where are corals most heat tolerant?
# Get mean and sd of PAMchange across all sites for each treatment
# mutate adds onto a df (vs summarise which makes one row for each group)
df2 <- group_by(df, Treatment) %>%
  mutate(temp_mean=mean(geno_mean, na.rm = TRUE),temp_sd=sd(geno_mean, na.rm = TRUE)) %>% 
  filter(geno_mean > (temp_mean + 1.5*temp_sd) )

df2sort <- arrange(df2,Site,Treatment,desc(Geno))

#not sure if this is actually using the right mean/sd for each temp -should be fewer than 8

ggplot(df2, aes(x = Site, fill= Site) ) +
  geom_bar() +
  facet_wrap(vars(Treatment)) +
  ggtitle("Most tolerant colonies by site (at each temperature)") 
```


# Run Stats on PAM data (modified  Barshis code)
```{r}
# make data into factors
pam.data$Geno <- as.factor(pam.data$Geno)
pam.data$Site <-  as.factor(pam.data$Site)
pam.data$Replicate <- as.factor(pam.data$Replicate)
pam.data$Treatment <- as.factor(pam.data$Treatment)
```

```{r mixed effects model PAM}
# Mixed Effects model - random effect of replicate
# Since the model is completely balanced the SS type I, II or III will provide the same result
pam.cbass.model <- lmer(PAMChange ~ Site * Treatment + (1|Replicate), data = pam.data)
anova(pam.cbass.model) # anova from lmerTest ANOVA SS Type III with ddf="Satterthwaite"
ranova(pam.cbass.model) # anova-like table for random effects
```

```{r model diagnostics PAM}
# Model fitting and assumptions diagnostic 
plot(PAMChange ~ interaction(Site,Treatment,Replicate), data = pam.data) # Box-plot homogeinity of variance
leveneTest(PAMChange ~ Site * Treatment * Replicate, data=pam.data) # formal statistical test for homogeinity of variance (not recommended due the small sample size)

plot(pam.cbass.model) # Residual vs Fitted values
qqnorm(resid(pam.cbass.model)); qqline(resid(pam.cbass.model)) # qq plot to check for normal distribution of residuals
hist(resid(pam.cbass.model)) # histogram of residuals to check for normal distribution of residuals
shapiro.test(resid(pam.cbass.model)) # formal statistical test (not recommended due the small sample size)
```

```{r post hoc PAM}
# Compare reef sites within each temperature treatment
pam.cbass.emms.reef <- emmeans(pam.cbass.model, pairwise ~ Site | Treatment, weights = "proportional", adjust="none")
summary(pam.cbass.emms.reef$emmeans)

# P.value adjustment of the Bonferroni
rbind(pam.cbass.emms.reef$contrasts, adjust="bonferroni")
```

```{r r Violin by Site PAM}
# Create violin plot
ggplot(data=pam.data, 
  aes(x=Treatment, y=PAMChange, label= Treatment, fill=Site)) +
  geom_violin(scale="width",outlier.size = 0.3, alpha=0.8) +
  geom_boxplot(width=0.1, color="black", alpha=0.2) +
  facet_grid(Species ~ Site) +
  expand_limits(y = 0)+
  theme_bw() +
  theme(legend.position = "none") +
  xlab(label = "Temperature (degC)") + 
  ylab(label = "Delta Fv/Fm")+
  theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust=1)) 

#save boxplot
ggsave(paste0("/Users/loga8761/github/PAM/Figures/PAM_Violin_BySite_"
              , format(Sys.time(), "%Y-%m-%d_%H-%M")
              , ".png"), width = 7, height = 4) # save as .png
```

# Run Stats on Color data (modified  Barshis code)
```{r mixed effects model Color}
# Mixed Effects model - random effect of replicate
# Since the model is completely balanced the SS type I, II or III will provide the same result
pam.cbass.model <- lmer(Color_2 ~ Site * Treatment + (1|Replicate), data = pam.data)
anova(pam.cbass.model) # anova from lmerTest ANOVA SS Type III with ddf="Satterthwaite"
ranova(pam.cbass.model) # anova-like table for random effects
```

```{r model diagnostics Color}
# Model fitting and assumptions diagnostic 
plot(Color_2 ~ interaction(Site,Treatment,Replicate), data = pam.data) # Box-plot homogeinity of variance
leveneTest(Color_2 ~ Site * Treatment * Replicate, data=pam.data) # formal statistical test for homogeinity of variance (not recommended due the small sample size)

plot(pam.cbass.model) # Residual vs Fitted values
qqnorm(resid(pam.cbass.model)); qqline(resid(pam.cbass.model)) # qq plot to check for normal distribution of residuals
hist(resid(pam.cbass.model)) # histogram of residuals to check for normal distribution of residuals
shapiro.test(resid(pam.cbass.model)) # formal statistical test (not recommended due the small sample size)
```

```{post hoc Color}
# comparing between reef sites within each temperature treatment
pam.cbass.emms.reef <- emmeans(pam.cbass.model, pairwise ~ Site | Treatment, weights = "proportional", adjust="none")
summary(pam.cbass.emms.reef$emmeans)

# P.value adjustment of the Bonferroni
rbind(pam.cbass.emms.reef$contrasts, adjust="bonferroni")
```

# Plot - Violin Plot for Color Change
```{r Violin Color,echo=TRUE}
# make violin plot
ggplot(pam.data,aes(x = Site, y = Color_2, fill = Site) )+
  geom_violin(scale="width",outlier.size = 0.3, alpha=0.8) +
  geom_boxplot(width=0.1, color="black", alpha=0.2) +
  facet_grid(Species ~ Treatment) +
  ylab("Final Color Score") +
  theme_bw() +
  scale_x_discrete(labels= SiteLabels) +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust=1)) 

# Save plot  
  ggsave(paste0("/Users/loga8761/github/PAM/Figures/Color2_Violin_by_site_"
              , format(Sys.time(), "%Y-%m-%d_%H-%M")
              , ".png"), width = 7, height = 4) # save as .png
```
# Violin plot by SITE 
```{r Violin by Site Color}
# Create violin plot
ggplot(data=pam.data, 
  aes(x=Treatment, y=Color_2, label= Treatment, fill=Site)) +
  geom_violin(scale="width",outlier.size = 0.3, alpha=0.8) +
  geom_boxplot(width=0.1, color="black", alpha=0.2) +
  facet_grid(Species ~ Site) +
  expand_limits(y = 0)+
  theme_bw() +
  theme(legend.position = "none") +
  xlab(label = "Temperature (degC)") + 
  ylab(label = "Final Color Score")+
  theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust=1)) 

#save boxplot
ggsave(paste0("/Users/loga8761/github/PAM/Figures/Color_Violin_BySite_"
              , format(Sys.time(), "%Y-%m-%d_%H-%M")
              , ".png"), width = 7, height = 4) # save as .png
```