---
title: "GLPS_stats_forCOAST"
author: "Cheryl Logan"
date: "9/11/2020"
output: html_document
---

## Load libraries
```{r setup, include = FALSE}
#library(thermPerf)
#install.packages("ggrepel")
library(tidyverse)
library(ggplot2)
library(RColorBrewer)
library(ggpubr)
library(ggrepel)
library(nlme)
library(tidyverse)
library(ggpubr)
library(rstatix)
library(forcats)
library(ggthemes)
library(lme4)
library(lmerTest)
library(emmeans)
library(car)
```

## Load data (from "Galapagos_PAM_allsites.Rmd")
```{r load data, include = FALSE}
PAM_all <- read_csv("/Users/loga8761/github/PAM/Galapagos_all.csv")
PAM_cold <- read_csv("/Users/loga8761/github/PAM/Galapagos_cold.csv")
PAM_heat <- read_csv("/Users/loga8761/github/PAM/Galapagos_heat.csv")
```

## R Stats for Pocillopora

```{r}
## PAM Pocillopora ANOVA model fitting & Tukey HSD
# https://www.datanovia.com/en/lessons/anova-in-r/#three-way-independent-anova
# for COAST

PAM_all_poc <- PAM_all %>%
  filter(Species == "Pocillopora_sp") 

```

ANOVA stats
```{r}
#Load the data and inspect one random row by group combinations:
PAM_all_poc %>% sample_n_by(Treatment, Site, Replicate, size = 1)

#Compute the mean and the standard deviation (SD) of PAM by groups:
PAM_all_poc %>%
  group_by(Treatment, Site, Replicate) %>%
  get_summary_stats(PAM, type = "mean_sd")

#Create a box plot of PAM by treatment
#note Isabela rep needs to be removed
bxp <- ggboxplot(
  PAM_all_poc, x = "Treatment", y = "PAM", 
  color = "Replicate", palette = "jco", facet.by = "Site"
  )
bxp

#ANOVA
res.aov <- PAM_all_poc %>% anova_test(PAM ~ Treatment*Site*Replicate)
res.aov

#Pocillopora
mod_anova_all <- aov(PAMChange ~ factor(Treatment)*Site*Replicate, data = PAM_all_poc)
summary(mod_anova_all)
TukeyHSD(mod_anova_all)

```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}


p <- PAM_all_poc %>%
  mutate(Site = fct_relevel(Site, 
            "Darwin", "Wolf2", "Punta_Pitt", 
            "Espanola", "Isabela")) %>%
  ggplot( aes(x=Site, y=PAMChange)) +
    geom_bar(stat="identity") +
    xlab("")
p

```


Plot used for 2020 COAST Proposedl
```{r,echo=TRUE}
#Labels for re-ordering site names
SiteLabels <- c("Darwin", "Wolf","San Cristobal", "Espanola", "Isabela","SantaCruz", "Floreana")

#re-order site names
p <- PAM_all %>%
  mutate(Site = fct_relevel(Site, 
            "Darwin", "Wolf2", "Punta_Pitt", 
            "Espanola", "Isabela"))  %>%
  filter(Site != "AcademyBay"& Site != "Floreana") %>% # Remove sites with data issues
  filter(Treatment != "14"& Site != "31") # Remove 14C and 31C temps since only used at Floreana
  
ggplot(p,aes(x = Site, y = PAMChange, fill = Site) )+
  geom_violin(scale="width",outlier.size = 0.3, alpha=0.8) +
  #geom_point(color = "black", size =.3, alpha = 0.8) +
  #geom_text(aes(label=Geno)) +
  #geom_violin(alpha=0.8) +
  geom_boxplot(width=0.1, color="black", alpha=0.2) +
  
  #stat_summary(fun.y="mean",geom="crossbar", mapping=aes(ymin=..y.., ymax=..y..), 
     #width=.5, position=position_dodge(),show.legend = FALSE) + 
  #facet_wrap(Species ~ Treatment) +
  facet_grid(Species ~ Treatment) +
  ylab("Change in Fv/Fm") +
  #xlab("Pocillopora by Site") +
  theme_bw() +
  #ylim(c(.5,-.45)) +
  scale_x_discrete(labels= SiteLabels) +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust=1)) +
  ggsave("Violin_by_site.png", width = 7, height = 4) 
  
  p

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

Create a plot with fraction of most heat- and cold- tolerant corals by site

```{r}
# For Pocillopora
PAM_all_poc_clean <- PAM_all_poc %>%
  filter(Site != "AcademyBay"& Site != "Floreana") %>%
  filter(Species == "Pocillopora_sp") %>%
  select(PAM, PAMChange,Site, Species,Treatment,Geno,Replicate)
  
# percent difference = abs((Value1 - Value2)/Value1)*100
df <- group_by(PAM_all_poc_clean,Site, Species,Treatment,Geno) %>%
  arrange(Site,Treatment,desc(Geno)) %>% # sort so replicates are next to each other
  mutate(geno_mean = mean(PAMChange, na.rm = TRUE)) %>% # genotype mean
  mutate(pct_diff = abs((PAMChange[1] - PAMChange[2])/PAMChange[1])*100, na.rm = TRUE)  
  

    
# Where are corals most heat tolerant?
# Get mean and sd of PAMchange across all sites for each treatment
# mutate adds onto a df (vs summarise which makes one row for each group)
df2 <- group_by(df, Treatment) %>%
  mutate(temp_mean=mean(geno_mean, na.rm = TRUE),temp_sd=sd(geno_mean, na.rm = TRUE)) %>% 
  filter(geno_mean > (temp_mean + 1.5*temp_sd) )

df2sort <- arrange(df2,Site,Treatment,desc(Geno))

#not sure if this is actually using the right mean/sd for each temp -should be fewer than 8

ggplot(df2, aes(x = Site, fill= Site) ) +
  geom_bar() +
  facet_wrap(vars(Treatment)) +
  ggtitle("Most tolerant colonies by site (at each temperature)") 


```


Run Stats with Barshis code
```{r}
#start with heat stress data only

pam.data <- PAM_all_poc_clean %>%
  filter(Treatment == "26" | Treatment == "32.5"| Treatment == "34"| Treatment == "35.5") 

pam.data$Geno <- as.factor(pam.data$Geno)
pam.data$Site <-  as.factor(pam.data$Site)
pam.data$Replicate <- as.factor(pam.data$Replicate)
pam.data$Treatment <- as.factor(pam.data$Treatment)
```


```{r}
# Mixed Effects model - random effect of replicate
# Since the model is completely balanced the SS type I, II or III will provide the same result
pam.cbass.model <- lmer(PAMChange ~ Site * Treatment + (1|Replicate), data = pam.data)
anova(pam.cbass.model) # anova from lmerTest ANOVA SS Type III with ddf="Satterthwaite"
ranova(pam.cbass.model) # anova-like table for random effects

```

```{r}
# Model fitting and assumptions diagnostic 
plot(PAMChange ~ interaction(Site,Treatment,Replicate), data = pam.data) # Box-plot homogeinity of variance
leveneTest(PAMChange ~ Site * Treatment * Replicate, data=pam.data) # formal statistical test for homogeinity of variance (not recommended due the small sample size)

plot(pam.cbass.model) # Residual vs Fitted values
qqnorm(resid(pam.cbass.model)); qqline(resid(pam.cbass.model)) # qq plot to check for normal distribution of residuals
hist(resid(pam.cbass.model)) # histogram of residuals to check for normal distribution of residuals
shapiro.test(resid(pam.cbass.model)) # formal statistical test (not recommended due the small sample size)
```

```{r}

# comparing between reef sites within each temperature treatment
pam.cbass.emms.reef <- emmeans(pam.cbass.model, pairwise ~ Site | Treatment, weights = "proportional", adjust="none")
summary(pam.cbass.emms.reef$emmeans)

# P.value adjustment of the Bonferroni
rbind(pam.cbass.emms.reef$contrasts, adjust="bonferroni")
```
```{r}
### BOX plots (Fix)
#pam <- ggplot(data=df, 
ggplot(data=pam.data, 
    aes(x=Treatment, y=PAMChange, label= Treatment, fill=Site)) +
  #scale_fill_manual(values = c ("#56B4E9", "#E69F00"), name = "Reef site") +
  stat_boxplot(geom ='errorbar', width = 0.7, lwd=0.7)+
  geom_boxplot(width=0.7, lwd=0.7, fatten=1) +
  expand_limits(y = 0)+
  facet_grid(~Site, space = "free", scales = "free")+ #this can also be used but rows and columns have to be specified -> facet_grid(. ~ experiment)
  theme_bw() +
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf, colour = "black", size=2) +

  theme(line= element_line(size = 1),
        axis.line = element_line(colour = "grey20"),
        axis.ticks.length = unit(0.2 , "cm"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank(),
        strip.background = element_blank(), 
        strip.text.x = element_text(color = "grey20", size = 20, angle = 0, hjust = 0, vjust = 0.5, face = "plain"),
        panel.spacing = unit(3, "lines"))+
      xlab(label = "Temperature treatment") + 
  ylab(label = "Delta photosynthetic efficiency (Fv/Fm)")+
  theme(axis.text.x = element_text(color = "grey20", size = 20, angle = 0, hjust = .5, vjust = .5, face = "plain"),
        axis.text.y = element_text(color = "grey20", size = 20, angle = 0, hjust = .5, vjust = .5, face = "plain"),  
        axis.title.x = element_text(color = "grey20", size = 20, angle = 0, hjust = .5, vjust = 0, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 20, angle = 90, hjust = .5, vjust = .5, face = "plain"),
        legend.title = element_text(colour="grey20", size=20, face="bold"),
        legend.text = element_text(colour="grey20", size=18, face="plain"),
        legend.position="bottom")

ggsave("Figures/PAMdata_Heat_POC_09062021.pdf", width = 10, height = 6)


```

