---
title: "CBASS_DRC_2022"
author: "Cheryl Logan"
date: "6/3/2022"
output: html_document
params:
  Species: Pocillopora_sp
  RampType: Heat
  printcode: TRUE
  data: /Users/loga8761/github/PAM/Data/Galapagos_all.csv
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load libraries
```{r setup, include = FALSE}
library(tidyverse)
library(ggpubr)
library(broom)
library(ggplot2) # v. 3.3.5
library(purrr) # v. 0.3.4
library(drc)
library(nls.multstart)
library(nlme)
library(lmerTest)
library(emmeans)
library(sjPlot)
library(drc)
library(Rmisc)
library(reshape2)
library(scales)
library(ggthemes)
library(ggpubr)
library(nlstools)

theme_set(theme_minimal())

```

# Load data 
```{r load.data, include = FALSE}
# from "Galapagos_PAM_allsites.Rmd"
# Remove sites with odd data or different temp profiles
PAM_all <- read_csv("/Users/loga8761/github/PAM/Data/Galapagos_all.csv") %>%
  filter(Site != "AcademyBay"& Site != "Floreana") %>%
  mutate(Site = fct_relevel(Site, 
            "Darwin", "Wolf2", "Punta_Pitt", 
            "Espanola", "Isabela"))  

pam.data <- PAM_all %>%
  filter(Species == params$Species) %>%
  filter(RampType == params$RampType) 

pam.data <- 
  pam.data %>% 
  mutate_at("Site", str_replace, "Pocillopora_sp", "POC") %>% 
  mutate_at("Site", str_replace, "Wolf2", "Wolf") %>% 
  mutate_at("Site", str_replace, "Punta_Pitt", "San_Cristobal")

#Labels for re-ordering site names (warm -> cold)
SiteLabels <- c("Darwin", "Wolf","San Cristobal", "Espanola", "Isabela")
#order sites Warmest to Coolest
pam.data$Site<-as.factor(pam.data$Site)
levels(pam.data$Site)
pam.data$Site = factor(pam.data$Site,levels(pam.data$Site)[c(1,5,4,2,3)]) 
# correct N to S order is: "Darwin", "Wolf","San Cristobal", "Espanola", "Isabela"
```

# Make data into factors
```{r load.data, include = FALSE}
pam.data$Site<-as.factor(pam.data$Site)
pam.data$Species<-as.factor(pam.data$Species)
pam.data$SiteSpecies<-as.factor(pam.data$Genobysite)
pam.data$Geno<-as.factor(pam.data$Geno)
pam.data$Replicate<-as.factor(pam.data$Replicate)
str(pam.data)

Pav_data<-subset(pam.data, Species == 'Pavona_clavus')
Poc_data<-subset(pam.data, Species == 'Pocillopora_sp')
```

# POCILLOPORA Dose-response curve model fitting - HEAT STRESS

#First, use mselect to find best curve fit
#m1<-drm(PAM ~ Treatment, data = Poc_data[Poc_data$Site=="Isabela",], fct = LL.3())
#mselect(m1, list(LL.5(), LL.4(), LL.3(), LL.2()), linreg=TRUE, icfct=AIC)
# Quad and LL.3 have lowest AIC values for heat stress; 
  
```{r poc drc isabela, include = FALSE}
#### Poc - ISABELA ####
      
#Run population-level fit using LL.3 (Based on Cunning et al. 2021) for heat stress only
Poc_Isabela_pop <- drm(PAM ~ Treatment, data = Poc_data[Poc_data$Site=="Isabela",], fct = LL.3(names = c("hill", "max", "ED50")),
      upperl = c(120, 0.72, 38),
      lowerl = c(10, 0.55, 30))
summary(Poc_Isabela_pop)
plot(Poc_Isabela_pop,broken = TRUE,xlim=c(26,40)  )

#Run individual genotype fits for heat stress only
Poc_Isabela <- drm(PAM ~ Treatment, data = Poc_data[Poc_data$Site=="Isabela",], curveid=Geno, fct = LL.3(names = c("hill", "max", "ED50")),
      upperl = c(120, 0.72, 38),
      lowerl = c(10, 0.55, 30))
summary(Poc_Isabela)
plot(Poc_Isabela,broken = TRUE,xlim=c(26,40)  )


#extract coeffs by geno, then compute 95% CIs - try ED50, or ED25 if temperatures weren't high enough
Poc_Isabela_genocoeffs_50<-data.frame(ED(Poc_Isabela, c(50)))
Poc_Isabela_coeff_mean<-mean(Poc_Isabela_genocoeffs_50$Estimate)
Poc_Isabela_coeff_mean

Poc_Isabela_summary<-data.frame(CI(Poc_Isabela_genocoeffs_50$Estimate, ci=0.95))
Poc_Isabela_coeff_lower<-Poc_Isabela_summary[3,]
Poc_Isabela_coeff_upper<-Poc_Isabela_summary[1,]
```

```{r poc drc espanola, include = FALSE}
#### Poc - ESPANOLA ####
      
#Run population-level fit using LL.3 (Based on Cunning et al. 2021) for heat stress only
Poc_Espanola_pop <- drm(PAM ~ Treatment, data = Poc_data[Poc_data$Site=="Espanola",], fct = LL.3(names = c("hill", "max", "ED50")),
      upperl = c(120, 0.72, 38),
      lowerl = c(10, 0.55, 30))
summary(Poc_Espanola_pop)
plot(Poc_Espanola_pop)

#Run individual genotype fits for heat stress only
Poc_Espanola <- drm(PAM ~ Treatment, data = Poc_data[Poc_data$Site=="Espanola",], curveid=Geno, fct = LL.3(names = c("hill", "max", "ED50")),
      upperl = c(120, 0.72, 38),
      lowerl = c(10, 0.55, 30))
summary(Poc_Espanola)
plot(Poc_Espanola)

#extract coeffs by geno, then compute 95% CIs - try ED50, or ED25 if temperatures weren't high enough
Poc_Espanola_genocoeffs_50<-data.frame(ED(Poc_Espanola, c(50)))
Poc_Espanola_coeff_mean<-mean(Poc_Espanola_genocoeffs_50$Estimate)
Poc_Espanola_coeff_mean

Poc_Espanola_summary<-data.frame(CI(Poc_Espanola_genocoeffs_50$Estimate, ci=0.95))
Poc_Espanola_coeff_lower<-Poc_Espanola_summary[3,]
Poc_Espanola_coeff_upper<-Poc_Espanola_summary[1,]
```

```{r poc drc san cristobal, include = FALSE}
#### Poc - SAN CRISOTOBAL ####
      
#Run population-level fit using LL.3 (Based on Cunning et al. 2021) for heat stress only
Poc_San_Cristobal_pop <- drm(PAM ~ Treatment, data = Poc_data[Poc_data$Site=="San_Cristobal",], fct = LL.3(names = c("hill", "max", "ED50")),
      upperl = c(120, 0.72, 38),
      lowerl = c(10, 0.55, 30))
summary(Poc_San_Cristobal_pop)
plot(Poc_San_Cristobal_pop)

#Run individual genotype fits for heat stress only
Poc_San_Cristobal <- drm(PAM ~ Treatment, data = Poc_data[Poc_data$Site=="San_Cristobal",], curveid=Geno, fct = LL.3(names = c("hill", "max", "ED50")),
      upperl = c(120, 0.72, 38),
      lowerl = c(10, 0.55, 30))
summary(Poc_San_Cristobal)
plot(Poc_San_Cristobal)

#extract coeffs by geno, then compute 95% CIs - try ED50, or ED25 if temperatures weren't high enough
Poc_San_Cristobal_genocoeffs_50<-data.frame(ED(Poc_San_Cristobal, c(50)))
Poc_San_Cristobal_coeff_mean<-mean(Poc_San_Cristobal_genocoeffs_50$Estimate)
Poc_San_Cristobal_coeff_mean

Poc_San_Cristobal_summary<-data.frame(CI(Poc_San_Cristobal_genocoeffs_50$Estimate, ci=0.95))
Poc_San_Cristobal_coeff_lower<-Poc_San_Cristobal_summary[3,]
Poc_San_Cristobal_coeff_upper<-Poc_San_Cristobal_summary[1,]
```

```{r poc drc wolf, include = FALSE}
#### Poc - WOLF ####
      
#Run population-level fit using LL.3 (Based on Cunning et al. 2021) for heat stress only
Poc_Wolf_pop <- drm(PAM ~ Treatment, data = Poc_data[Poc_data$Site=="Wolf",], fct = LL.3(names = c("hill", "max", "ED50")),
      upperl = c(120, 0.72, 38),
      lowerl = c(10, 0.55, 30))
summary(Poc_Wolf_pop)
plot(Poc_Wolf_pop)

#Run individual genotype fits for heat stress only
Poc_Wolf <- drm(PAM ~ Treatment, data = Poc_data[Poc_data$Site=="Wolf",], curveid=Geno, fct = LL.3(names = c("hill", "max", "ED50")),
      upperl = c(120, 0.72, 38),
      lowerl = c(10, 0.55, 30))
summary(Poc_Wolf)
plot(Poc_Wolf)

#extract coeffs by geno, then compute 95% CIs - try ED50, or ED25 if temperatures weren't high enough
Poc_Wolf_genocoeffs_50<-data.frame(ED(Poc_Wolf, c(50)))
Poc_Wolf_coeff_mean<-mean(Poc_Wolf_genocoeffs_50$Estimate)
Poc_Wolf_coeff_mean

Poc_Wolf_summary<-data.frame(CI(Poc_Wolf_genocoeffs_50$Estimate, ci=0.95))
Poc_Wolf_coeff_lower<-Poc_Wolf_summary[3,]
Poc_Wolf_coeff_upper<-Poc_Wolf_summary[1,]
```

```{r poc darwin, include = FALSE}
#### Poc - DARWIN ####
      
#Run population-level fit using LL.3 (Based on Cunning et al. 2021) for heat stress only
Poc_Darwin_pop <- drm(PAM ~ Treatment, data = Poc_data[Poc_data$Site=="Darwin",], fct = LL.3(names = c("hill", "max", "ED50")),
      upperl = c(120, 0.72, 38),
      lowerl = c(10, 0.55, 30))
summary(Poc_Darwin_pop)
plot(Poc_Darwin_pop)

#Run individual genotype fits for heat stress only
Poc_Darwin <- drm(PAM ~ Treatment, data = Poc_data[Poc_data$Site=="Darwin",], curveid=Geno, fct = LL.3(names = c("hill", "max", "ED50")),
      upperl = c(120, 0.72, 38),
      lowerl = c(10, 0.55, 30))
summary(Poc_Darwin)
plot(Poc_Darwin)

#extract coeffs by geno, then compute 95% CIs - try ED50, or ED25 if temperatures weren't high enough
Poc_Darwin_genocoeffs_50<-data.frame(ED(Poc_Darwin, c(50)))
Poc_Darwin_coeff_mean<-mean(Poc_Darwin_genocoeffs_50$Estimate)
Poc_Darwin_coeff_mean

Poc_Darwin_summary<-data.frame(CI(Poc_Darwin_genocoeffs_50$Estimate, ci=0.95))
Poc_Darwin_coeff_lower<-Poc_Darwin_summary[3,]
Poc_Darwin_coeff_upper<-Poc_Darwin_summary[1,]
```

Plots
```{r}
par(mfrow=c(3,2))
plot(Poc_Darwin_pop, broken = TRUE,xlim=c(26,40)); title(main = "Darwin ")  
plot(Poc_Wolf_pop, broken = TRUE,xlim=c(26,40)); title(main = "Wolf ")  
plot(Poc_San_Cristobal_pop, broken = TRUE,xlim=c(26,40)); title(main = "San Cristobal ")
plot(Poc_Espanola_pop, broken = TRUE,xlim=c(26,40)); title(main = "Espanola ")
plot(Poc_Isabela_pop, broken = TRUE,xlim=c(26,40)); title(main = "Isabela ")
```

############################################################################################
#### Combine genotpye-ED50s into dataframe for statistical analysis ####
############################################################################################
```{r combine ED25s, include = FALSE}
Poc_Geno_ED50s <- data.frame(cbind(Poc_Darwin_genocoeffs_50[,1],Poc_Wolf_genocoeffs_50[,1],Poc_San_Cristobal_genocoeffs_50[,1],
                    Poc_Espanola_genocoeffs_50[,1], Poc_Isabela_genocoeffs_50[,1]))

Poc_Geno_ED50s<-Poc_Geno_ED50s %>% 
  dplyr::rename(Darwin= X1,
         Wolf=X2,
         San_Cristobal=X3,
         Espanola=X4,
         Isabela=X5)

Poc_Geno_ED50s$Geno<-as.factor(1:nrow(Poc_Geno_ED50s))
str(Poc_Geno_ED50s)

Poc_Geno_ED50s_long<-melt(Poc_Geno_ED50s, id="Geno")

Poc_Geno_ED50s_long<-Poc_Geno_ED50s_long %>% 
  dplyr::rename(Site= variable,
         ED50=value)

Poc_Geno_ED50s_long$Species<-rep('Pocillopora_sp')
Poc_Geno_ED50s_long$MMM<-rep(c('26','26','26','26','26'),each=8)

Poc_Geno_ED50s_long$MMM<-as.numeric(Poc_Geno_ED50s_long$MMM)
Poc_Geno_ED50s_long['Relative_Threshold'] = Poc_Geno_ED50s_long['ED50'] - Poc_Geno_ED50s_long['MMM']

Poc_ED50_mod<-aov(ED50 ~ Site, Poc_Geno_ED50s_long)
summary(Poc_ED50_mod)
TukeyHSD(Poc_ED50_mod)
```


############################################################################################
#### Combine ED50 data plus predict curves from models for plotting ####
############################################################################################
```{r prep plot, include = FALSE}
Poc_coeff_means<-data.frame(Poc_Darwin_coeff_mean, Poc_Wolf_coeff_mean, Poc_San_Cristobal_coeff_mean, Poc_Espanola_coeff_mean, Poc_Isabela_coeff_mean)
Poc_coeff_lowers<-data.frame(Poc_Darwin_coeff_lower, Poc_Wolf_coeff_lower, Poc_San_Cristobal_coeff_lower, Poc_Espanola_coeff_lower, Poc_Isabela_coeff_lower)
Poc_coeff_uppers<-data.frame(Poc_Darwin_coeff_upper, Poc_Wolf_coeff_upper, Poc_San_Cristobal_coeff_upper, Poc_Espanola_coeff_upper, Poc_Isabela_coeff_upper)

Poc_Darwin_preddata = data.frame(temp = seq(30,39, length.out = 100))
Poc_Darwin_pred = as.data.frame(predict(Poc_Darwin_pop, newdata = Poc_Darwin_preddata, interval = 'confidence'))
Poc_Darwin_preddata = data.frame(Poc_Darwin_preddata, fvfm = Poc_Darwin_pred$Prediction, Lower = Poc_Darwin_pred$Lower, Upper = Poc_Darwin_pred$Upper)

Poc_Wolf_preddata = data.frame(temp = seq(30,39, length.out = 100))
Poc_Wolf_pred = as.data.frame(predict(Poc_Wolf_pop, newdata = Poc_Wolf_preddata, interval = 'confidence'))
Poc_Wolf_preddata = data.frame(Poc_Wolf_preddata, fvfm = Poc_Wolf_pred$Prediction, Lower = Poc_Wolf_pred$Lower, Upper = Poc_Wolf_pred$Upper)

Poc_San_Cristobal_preddata = data.frame(temp = seq(30,39, length.out = 100))
Poc_San_Cristobal_pred = as.data.frame(predict(Poc_San_Cristobal_pop, newdata = Poc_San_Cristobal_preddata, interval = 'confidence'))
Poc_San_Cristobal_preddata = data.frame(Poc_San_Cristobal_preddata, fvfm = Poc_San_Cristobal_pred$Prediction, Lower = Poc_San_Cristobal_pred$Lower, Upper = Poc_San_Cristobal_pred$Upper)

Poc_Espanola_preddata = data.frame(temp = seq(30,39, length.out = 100))
Poc_Espanola_pred = as.data.frame(predict(Poc_Espanola_pop, newdata = Poc_Espanola_preddata, interval = 'confidence'))
Poc_Espanola_preddata = data.frame(Poc_Espanola_preddata, fvfm = Poc_Espanola_pred$Prediction, Lower = Poc_Espanola_pred$Lower, Upper = Poc_Espanola_pred$Upper)

Poc_Isabela_preddata = data.frame(temp = seq(30,39, length.out = 100))
Poc_Isabela_pred = as.data.frame(predict(Poc_Isabela_pop, newdata = Poc_Isabela_preddata, interval = 'confidence'))
Poc_Isabela_preddata = data.frame(Poc_Isabela_preddata, fvfm = Poc_Isabela_pred$Prediction, Lower = Poc_Isabela_pred$Lower, Upper = Poc_Isabela_pred$Upper)
```

```{r}
#### PLOT Pocollopora ####
Poc_data$Treatment<-as.character(Poc_data$Treatment)
Poc_data$Treatment<-as.numeric(Poc_data$Treatment)
#need to change y-axis for FvFm data

Poc_plot<- ggplot() +
  geom_jitter(data = Poc_data, alpha = 0.2, aes(x = Treatment, y = PAM, color = Site), size = 1, width = 0.25) +
  scale_x_continuous(limits=c(29,40), breaks=c(30,32,34,36,38)) +
  scale_y_continuous(limits=c(-0.01, 0.75), breaks=c(0, 0.2, 0.4, 0.6)) +
  
  geom_line(data = Poc_Darwin_preddata, aes(x = temp, y = fvfm), color = 'royalblue2', show.legend = FALSE) +
  geom_ribbon(data = Poc_Darwin_preddata, aes(x = temp, ymin=Lower, ymax=Upper), color = 'royalblue2', linetype=2, alpha = 0.2) +
  geom_vline(data = Poc_coeff_means, aes(xintercept = Poc_Darwin_coeff_mean), color = 'royalblue2', show.legend = FALSE) +
  annotate("rect", xmin=Poc_coeff_lowers$Poc_Darwin_coeff_lower, xmax=Poc_coeff_uppers$Poc_Darwin_coeff_upper, ymin=-Inf, ymax=Inf, fill= 'royalblue2',  alpha = 0.1) +
  geom_text(data = Poc_coeff_means, aes(label=round(Poc_Darwin_coeff_mean, digits = 2)), x = 31, y = 0.3, show.legend = FALSE, color = 'royalblue2') +

  geom_line(data = Poc_Wolf_preddata, aes(x = temp, y = fvfm), color = 'darkgoldenrod1', show.legend = FALSE) +
  #geom_ribbon(data = Poc_Wolf_preddata, aes(x = temp, ymin=Lower, ymax=Upper), color = 'darkgoldenrod1', linetype=2, alpha = 0.2) +
  #geom_vline(data = Poc_coeff_means, aes(xintercept = Poc_Wolf_coeff_mean), color = 'darkgoldenrod1', show.legend = FALSE) +
  annotate("rect", xmin=Poc_coeff_lowers$Poc_Wolf_coeff_lower, xmax=Poc_coeff_uppers$Poc_Wolf_coeff_upper, ymin=-Inf, ymax=Inf, fill= 'darkgoldenrod1',  alpha = 0.1) +
  geom_text(data = Poc_coeff_means, aes(label=round(Poc_Wolf_coeff_mean, digits = 2)), x = 31, y = 0.25, show.legend = FALSE, color = 'darkgoldenrod1') +

  geom_line(data = Poc_San_Cristobal_preddata, aes(x = temp, y = fvfm), color = 'darkorange1', show.legend = FALSE) +
  #geom_ribbon(data = Poc_San_Cristobal_preddata, aes(x = temp, ymin=Lower, ymax=Upper), color = 'darkorange1', linetype=2, alpha = 0.2) +
  #geom_vline(data = Poc_coeff_means, aes(xintercept = Poc_San_Cristobal_coeff_mean), color = 'darkorange1', show.legend = FALSE) +
  annotate("rect", xmin=Poc_coeff_lowers$Poc_San_Cristobal_coeff_lower, xmax=Poc_coeff_uppers$Poc_San_Cristobal_coeff_upper, ymin=-Inf, ymax=Inf, fill= 'darkorange1',  alpha = 0.1) +
  geom_text(data = Poc_coeff_means, aes(label=round(Poc_San_Cristobal_coeff_mean, digits = 2)), x = 31, y = 0.20, show.legend = FALSE, color = 'darkorange1') +

  geom_line(data = Poc_Espanola_preddata, aes(x = temp, y = fvfm), color = 'red3', show.legend = FALSE) +
  #geom_ribbon(data = Poc_Espanola_preddata, aes(x = temp, ymin=Lower, ymax=Upper), color = 'red3', linetype=2, alpha = 0.2) +
  #geom_vline(data = Poc_coeff_means, aes(xintercept = Poc_Espanola_coeff_mean), color = 'red3', show.legend = FALSE) +
  annotate("rect", xmin=Poc_coeff_lowers$Poc_Espanola_coeff_lower, xmax=Poc_coeff_uppers$Poc_Espanola_coeff_upper, ymin=-Inf, ymax=Inf, fill= 'red3',  alpha = 0.1) +
  geom_text(data = Poc_coeff_means, aes(label=round(Poc_Espanola_coeff_mean, digits = 2)), x = 31, y = 0.15, show.legend = FALSE, color = 'red3') +
  
  geom_line(data = Poc_Isabela_preddata, aes(x = temp, y = fvfm), color = 'springgreen1', show.legend = FALSE) +
  #geom_ribbon(data = Poc_Isabela_preddata, aes(x = temp, ymin=Lower, ymax=Upper), color = 'springgreen1', linetype=2, alpha = 0.2) +
  #geom_vline(data = Poc_coeff_means, aes(xintercept = Poc_Isabela_coeff_mean), color = 'springgreen1', show.legend = FALSE) +
  annotate("rect", xmin=Poc_coeff_lowers$Poc_Isabela_coeff_lower, xmax=Poc_coeff_uppers$Poc_Isabela_coeff_upper, ymin=-Inf, ymax=Inf, fill= 'springgreen1',  alpha = 0.1) +
  geom_text(data = Poc_coeff_means, aes(label=round(Poc_Isabela_coeff_mean, digits = 2)), x = 31, y = 0.10, show.legend = FALSE, color = 'springgreen1') +
  
  ggtitle("Pocillopora sp. ED50") +
  scale_color_manual(values=c("royalblue2", "darkgoldenrod1", "darkorange1", "red3", "springgreen1")) +
  ylab("Fv/Fm") +
  xlab("Temperature (°C)") +
  theme_bw()
Poc_plot
```



# POCILLOPORA DRCs - COLD STRESS

```{r poc drc, include = FALSE}
#### Poc - ISABELA ####

Poc_data <- Poc_data %>% 
  filter(Site == "Darwin") %>%
  unite("GenoRep", Geno:Replicate, sep= "-",remove = FALSE) %>%
  dplyr::select(GenoRep, Geno, Replicate, Treatment, PAM) %>%
  arrange(GenoRep)

# to work with tidy data, need to tidy data
Poc_data$GenoRep<-as.factor(Poc_data$GenoRep)
Poc_data <- as_tibble(Poc_data)
Poc_data

# To fit multiple models, use a nest-map-unnest workflow. 
# We start by nesting our data based on the group of interest (Genotype or Replicate)
# nesting is similar to group_by
# Then  we can handle multiple regressions at once:
# 'map' will tidy each of the objects

regressions <- Poc_data %>%
  #nest(data = -GenoRep) %>% # nest by Replicate
  nest(data = -Geno) %>% # nest by Genotype
  mutate(
    fit = map(data, ~ lm(PAM ~ log(Treatment), data = .x)), # S3 list-col
    tidied = map(fit, tidy),
    glanced = map(fit, glance),
    augmented = map(fit, augment)
  ) 

#regressions %>% 
  unnest(tidied)

regressions %>% # most useful - gives R^2 and p-value for all models
  unnest(glanced)

#regressions %>% 
  unnest(augmented)
  
## can't figure out easy way to plot

# Set Up New Data To Predict On
# following https://dpananos.github.io/posts/2018/04/blog-post-8/
regressions_new = regressions %>%
		 mutate(newdata = map(data, ~ data_grid(.x, PAM = seq_range(PAM,10)) ) )

# Predict & Transform NOT WORKING
step.5 = regressions_new %>%
		 mutate(preds = map2(regressions, newdata, ~ exp(predict(.x, newdata = .y))) ) #Dont forget the exp!



```