---
title: "A report for `r params$Species` `r params$RampType`"
author: "Cheryl Logan"
date: "2/15/2022"
output: html_document
params:
  Species: Pocillopora_sp
  RampType: Cold
  printcode: TRUE
  data: /Users/loga8761/github/PAM/Galapagos_all.csv

---

Script adapted from Voolstra et al. 2020 GCB
https://doi.org/10.1111/gcb.15148
https://github.com/reefgenomics/CBASSvsCLASSIC

The script is used for change in Fv/Fm and Color Card change. Modify variable Var:
Var <- pam.data$ColorChange # Color Card Change
Var <- pam.data$Color_2 # Color Card Endpoint
Var <- pam.data$PAM_Change #Fv/Fm

# Load libraries
```{r setup, include = FALSE}
library(tidyverse)
library(ggplot2)
library(car)
library(lmerTest)
library(emmeans)
library(ggplot2)
library(ggpubr)
library(sjPlot)
library(sjmisc)
library(broom)
today.date <- Sys.Date()
```

# Load data 
```{r load.data, include = FALSE}
# from "Galapagos_PAM_allsites.Rmd"
# Remove sites with odd data or different temp profiles
PAM_all <- read_csv("/Users/loga8761/github/PAM/Data/Galapagos_all.csv") %>%
  filter(Site != "AcademyBay"& Site != "Floreana") %>%
  mutate(Site = fct_relevel(Site, 
            "Darwin", "Wolf2", "Punta_Pitt", 
            "Espanola", "Isabela"))  

pam.data <- PAM_all %>%
  filter(Species == params$Species) %>%
  filter(RampType == params$RampType) %>%                               # Replacing values
  mutate_at("Site",str_replace,"Wolf2", "Wolf") %>%                               # Replacing values
  mutate_at("Site",str_replace,"Punta_Pitt", "San Cristobal")
  
#Labels for re-ordering site names (warm -> cold)
sitelabs <- c("Darwin", "Wolf2", "Punta_Pitt", "Espanola", "Isabela")
names(sitelabs) <- c("Darwin", "Wolf","San Cristobal", "Espanola", "Isabela")
```

# Set Response variable
```{r load.var, include = FALSE}
Var <- pam.data$PAMChange     #Fv/Fm
#Var <- pam.data$ColorChange    # Color Card Change
#Var <- pam.data$Color_2         # Color Card Endpoint
```

# Subset data
```{r load data subset, include = TRUE}
#pam.data <- x

pam.data$Geno <- as.factor(pam.data$Geno)
pam.data$Genobysite <- as.factor(pam.data$Genobysite)
pam.data$Site <-  as.factor(pam.data$Site)
pam.data$Replicate <- as.factor(pam.data$Replicate)
pam.data$Treatment <- as.factor(pam.data$Treatment)
```

# Mixed Effects model 
```{r lme model, include = TRUE}
# random effect of replicate 
# Since the model is completely balanced the SS type I, II or III will provide the same result
pam.cbass.model <- lmer(Var ~ Site * Treatment + (1|Replicate), data = pam.data)
anova(pam.cbass.model) # anova from lmerTest ANOVA SS Type III with ddf="Satterthwaite"
ranova(pam.cbass.model) # anova-like table for random effects
```

# Model fitting and assumptions diagnostic
```{r fitting, include = TRUE}
plot(Var ~ interaction(Site,Treatment,Replicate), data = pam.data) # Box-plot homogeneity of variance
leveneTest(Var ~ Site * Treatment * Replicate, data=pam.data) # formal statistical test for homogeinity of variance (not recommended due the small sample size)
```

# Residual vs Fitted values
```{r residuals, include = TRUE}
plot(pam.cbass.model) # Residual vs Fitted values
qqnorm(resid(pam.cbass.model)); qqline(resid(pam.cbass.model)) # qq plot to check for normal distribution of residuals
hist(resid(pam.cbass.model)) # histogram of residuals to check for normal distribution of residuals
shapiro.test(resid(pam.cbass.model)) # formal statistical test (not recommended due the small sample size)
```

# Compare sites by temperature treatment
```{r compare , include = TRUE}
pam.cbass.emms.reef <- emmeans(pam.cbass.model, pairwise ~ Site | Treatment, weights = "proportional", adjust="none")
summary(pam.cbass.emms.reef$emmeans)
```

# P.value adjustment with Bonferroni
```{r, bonferroni, include = TRUE}
rbind(pam.cbass.emms.reef$contrasts, adjust="bonferroni")
```

# Plot data
```{r plot}

ggplot(data=pam.data, 
    aes(x=Treatment, y=Var, label= Treatment, fill=Site)) +
  #scale_fill_manual(values = c ("#56B4E9", "#E69F00"), name = "Reef site") +
  stat_boxplot(geom ='errorbar', width = 0.7, lwd=0.7)+
  geom_boxplot(width=0.7, lwd=0.7, fatten=1) +
  expand_limits(y = 0)+
  #facet_grid(~Site, space = "free", scales = "free")+ #this can also be used but rows and columns have to be specified -> facet_grid(. ~ experiment)
  facet_grid(~factor(Site, levels=c("Darwin", "Wolf","San Cristobal", "Espanola", "Isabela")))+
  theme_bw() +
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf, colour = "black", size=2)  + 
  ylab(" Fv/Fm") +
  ggtitle(paste0('PAM_', params$Species, '_', params$RampType))


# ggsave("ColorEndpoint_Cold_POC_09072021.pdf", width = 10, height = 6)
# save
filename=paste0('GenotypeTolerance_', params$Species, '_', params$RampType, '.','pdf')
ggsave(path= "/Users/loga8761/github/PAM/Figures", filename, width = 10, height = 6)
```

# Identify top 10 most/least tolerant colonies

```{r}
#Fit linear model for each genotype across all temperatures
pam.data$Treatment <- as.numeric(as.character(pam.data$Treatment))
pam.data <- pam.data %>% drop_na(PAMChange) # remove NAs
slopes<-pam.data %>% group_by(Genobysite) %>% do(tidy(lm(PAMChange ~ Treatment, .)))

strong <- slopes %>%
  ungroup() %>%
  filter(term=="Treatment") %>%   # Select slopes only
  arrange(abs(estimate)) %>%      # sort by absolute value of slope
  slice_head(n = 10) %>%
  mutate(tolerance="most_tolerant")
strong_corals <-right_join(pam.data, strong, by = "Genobysite")

weak <- slopes %>%
  ungroup() %>%
  filter(term=="Treatment") %>% 
  arrange(abs(estimate)) %>%
  slice_tail(n = 10) %>%
  mutate(tolerance="least_tolerant")
weak_corals <-right_join(pam.data, weak, by = "Genobysite")
          
weak.strong <-bind_rows(strong_corals, weak_corals)

```

Isabela and Darwin only
```{r}
#Fit linear model for each genotype across all temperatures
pam.data$Treatment <- as.numeric(as.character(pam.data$Treatment))
pam.data <- pam.data %>% drop_na(PAMChange) # remove NAs
pam.data <- pam.data %>% filter(Site=="Isabela" | Site == "Darwin") # remove NAs
slopes<-pam.data %>% group_by(Genobysite) %>% do(tidy(lm(PAMChange ~ Treatment, .)))

Isa.Dar.strong <- slopes %>%
  ungroup() %>%
  filter(term=="Treatment") %>%   # Select slopes only
  arrange(abs(estimate)) 

Isa.Dar.strong_corals <-right_join(pam.data, Isa.Dar.strong, by = "Genobysite")%>%    
  arrange(desc(abs(estimate)))   %>%   # sort by absolute value of slope
  group_by(Site) %>% mutate(rank = dense_rank(estimate))# sort by absolute value of slope
  #select(Genobysite, Site,estimate) %>%
  #unique()

ggplot(data=Isa.Dar.strong_corals, 
    aes(x=Treatment, y=PAMChange, label= Treatment, group=Genobysite, color=rank)) +
  #scale_fill_manual(values = c ("#56B4E9", "#E69F00"), name = "Reef site") +
  geom_point()+
  geom_smooth(method = "lm", se = FALSE)+
  facet_grid(~Site, space = "free", scales = "free") + #this can also be used but rows and columns have to be specified -> facet_grid(. ~ experiment) +
  # geom_text(
  #   label=rownames(Geno), 
  #   nudge_x = 0.25, nudge_y = 0.25, 
  #   check_overlap = T
  # ) +
  theme_bw() +
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf, colour = "black", size=2)+
  ggtitle(paste0('GenotypeTolerance_', params$Species, '_', params$RampType))

# save
filename=paste0('IsaDarGenotypeTolerance_',params$Species,'_', params$RampType,'_',today.date,'.','pdf')
ggsave(path="/Users/loga8761/github/PAM/Figures", filename, width = 10, height = 6)
filename=paste0('IsaDarStrong_',params$Species,'_', params$RampType,'_',today.date,'.','csv')
write_csv(Isa.Dar.strong, file = filename)
```

# Plot weakest & stronge genotypes
```{r}
ggplot(data=weak.strong, 
    aes(x=Treatment, y=PAMChange, label= Treatment, group=Genobysite, color=Site)) +
  #scale_fill_manual(values = c ("#56B4E9", "#E69F00"), name = "Reef site") +
  geom_point()+
  geom_smooth(method = "lm", se = FALSE)+
  facet_grid(~tolerance, space = "free", scales = "free")+ #this can also be used but rows and columns have to be specified -> facet_grid(. ~ experiment)
  theme_bw() +
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf, colour = "black", size=2)+
  ggtitle(paste0('GenotypeTolerance_', params$Species, '_', params$RampType))

# save
filename=paste0('GenotypeTolerance_',params$Species,'_', params$RampType,'_',today.date,'.','pdf')
ggsave(path="/Users/loga8761/github/PAM/Figures", filename, width = 10, height = 6)
```

# Plot weakest & stronge genotypes by site
```{r}
hist.weak.strong <- weak.strong %>%
  select(Genobysite, Site,estimate, tolerance) %>%
  unique()

hist.weak.strong$Site <- factor(hist.weak.strong$Site,  # Change ordering manually
                  levels = c("Darwin", "Wolf", "San Cristobal", "Espanola", "Isabela"))

ggplot(data=hist.weak.strong, 
  aes(x=Site , fill = tolerance)) +
  #scale_fill_manual(values = c ("#56B4E9", "#E69F00"), name = "Reef site") +
  geom_bar()+
  theme_bw() +
  ggtitle(paste0('Top10_Genotypes_', params$Species, '_', params$RampType)) 

#write_csv(hist.weak.strong, file = "data_output/surveys_complete.csv")

# save
filename=paste0('Top10_Genotypes_',params$Species,'_', params$RampType,'_',today.date,'.','pdf')
ggsave(path="/Users/loga8761/github/PAM/Figures", filename, width = 10, height = 6)
```

# Plot weakest & stronge genotypes for Isabela and Darwin only
```{r}
Isa.Dar.hist.weak.strong <- weak.strong %>%
  select(Genobysite, Site,estimate, tolerance) %>%
  unique()

Isa.Dar.hist.weak.strong$Site <- factor(hist.weak.strong$Site,  # Change ordering 
                  levels = c("Darwin", "Isabela"))

ggplot(data=Isa.Dar.hist.weak.strong, 
  aes(x=Site , fill = tolerance)) +
  #scale_fill_manual(values = c ("#56B4E9", "#E69F00"), name = "Reef site") +
  geom_bar()+
  theme_bw() +
  ggtitle(paste0('Top10_Genotypes_', params$Species, '_', params$RampType)) 

#write_csv(hist.weak.strong, file = "data_output/surveys_complete.csv")

# save
filename=paste0('Top10_Genotypes_',params$Species,'_', params$RampType,'_',today.date,'.','pdf')
ggsave(path="/Users/loga8761/github/PAM/Figures", filename, width = 10, height = 6)
```

