---
title: "rTPC_v2"
author: "Cheryl Logan"
date: '2022-06-14'
output: html_document
---

```{r setup, echo =FALSE}
# load packages
#remotes::install_github("padpadpadpad/rTPC")
library(rTPC)
library(nls.multstart)
library(broom)
library(tidyverse)
library(ggrepel)
library(MuMIn)

# write function to label ggplot2 panels
label_facets_num <- function(string){
  len <- length(string)
  string = paste('(', 1:len, ') ', string, sep = '')
  return(string)
}
today.date <- Sys.Date()
```

## Goal: Produce Thermal Performance Curves for GLPS coral data following the rTPC tutorials: 
https://padpadpadpad.github.io/rTPC/articles/rTPC.html

```{r get model names, echo=FALSE}
get_model_names()
```

# Load GLPS PAM data
```{r load data, echo=FALSE}
# Load data 

# from "Galapagos_PAM_allsites.Rmd"
# Remove sites with odd data or different temp profiles
PAM_all <- read_csv("/Users/loga8761/github/PAM/Data/Galapagos_all.csv") %>%
  filter(Site != "AcademyBay"& Site != "Floreana") %>%
  mutate(Site = fct_relevel(Site, 
            "Darwin", "Wolf2", "Punta_Pitt", 
            "Espanola", "Isabela"))  

pam.data <- PAM_all %>%
  filter(Species == "Pocillopora_sp") %>%
  #filter(RampType == params$RampType) %>%                               # Replacing values
  mutate_at("Site",str_replace,"Wolf2", "Wolf") %>%                               # Replacing values
  mutate_at("Site",str_replace,"Punta_Pitt", "San Cristobal") %>%  
  select(Geno,Treatment, PAM, Site)

pam.data$Geno <- as.numeric(pam.data$Geno)
pam.data$Site <-  as.factor(pam.data$Site)
pam.data$Treatment <- as.numeric(pam.data$Treatment)
  
#Labels for re-ordering site names (warm -> cold)
sitelabs <- c("Darwin", "Wolf2", "Punta_Pitt", "Espanola", "Isabela")
names(sitelabs) <- c("Darwin", "Wolf","San Cristobal", "Espanola", "Isabela")

# Get subset of data
Isabela<-pam.data %>% 
  filter(Site == "Isabela")  %>%
  arrange(Geno)

#rename columns to follow tutorial more easily
Isabela <- as.data.frame(Isabela)  %>%
  rename(temp = Treatment)  %>%
  rename(rate = PAM)  %>%
  na.omit()
```

# Select data from a single site and genotype to try
```{r pressure, echo=FALSE}
# load in sample data
data("chlorella_tpc")

# keep just a single curve
# d <- filter(chlorella_tpc, curve_id == 1) # sample data
d <- filter(Isabela, Geno == 5)

# plot data
ggplot(d, aes(temp, rate)) +
  geom_point() +
  theme_bw(base_size = 12) +
  labs(x = 'Temperature (ºC)',
       y = 'FvFm',
       title = 'Photosynthetic efficiency across temperature')
```

## Estimate starting values for TPC model

Estimate sensible start values (get_start_vals()), lower (get_lower_lims()) and upper (get_upper_lims()) limits. To demonstrate this, use Johnson-Lewin model (Johnson and Lewin, 1946)

```{r get start vals, echo=FALSE}
# get start values and fit model
start_vals <- get_start_vals(d$temp, d$rate, model_name = 'johnsonlewin_1946')
# fit model
mod <- suppressWarnings(
nls.multstart::nls_multstart(rate~johnsonlewin_1946(temp = temp, r0, e, eh, topt),
data = d,
iter = c(5,5,5,5),
start_lower = start_vals - 1,
start_upper = start_vals + 1,
lower = get_lower_lims(d$temp, d$rate, model_name = 'johnsonlewin_1946'),
upper = get_upper_lims(d$temp, d$rate, model_name = 'johnsonlewin_1946'),
supp_errors = 'Y',
convergence_count = FALSE))

# look at model fit
summary(mod)
```
Using nls_multstart, we will use a random-search/shotgun approach to fit the curve. Random start parameter values are picked from a uniform distribution between start_lower and start_upper for each parameter. If the best model is not improved upon (in terms of AIC score) for 100 new start parameter combinations, the function will return that model fit. This is controlled by convergence_count, if this is set to FALSE, nls_multstart() will try and fit all iterations.

# Get model predictions and plot TPC model on actual data

```{r plot preds, echo=FALSE}
# get predictions
preds <- data.frame(temp = seq(min(d$temp), max(d$temp), length.out = 100))
preds <- broom::augment(mod, newdata = preds)

# plot
ggplot(preds) +
geom_point(aes(temp, rate), d) +
geom_line(aes(temp, .fitted), col = 'blue') +
theme_bw()+
  labs(x = 'Temperature (ºC)',
       y = 'FvFm',
       title = 'Photosynthetic efficiency across temperature')
```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.


# Model selection 
1. Fit all models in rTPC and plot using 1 example curve from Darwin
2. Next, I manually removed models that obviously did not fit well at all:
boatman, briere2, joehnk, lrf, ratkowsky, sharpeschoolfull, sharpeschoollow, modified gaussian, kamykowski, gaussian, joehnk, delong, flinn, quadratic, weibull, thomas1
3. Run  models on other example curves from Isabela and Darwin

```{r fit many models, echo = FALSE}
# fit every model formulation in rTPC
d_fits <- nest(d, data = c(temp, rate)) %>%
  mutate(beta = map(data, ~nls_multstart(rate~beta_2012(temp = temp, a, b, c, d, e),
                        data = .x,
                        iter = c(6,6,6,6,6),
                        start_lower = get_start_vals(.x$temp, .x$rate, model_name = 'beta_2012') - 10,
                        start_upper = get_start_vals(.x$temp, .x$rate, model_name = 'beta_2012') + 10,
                        lower = get_lower_lims(.x$temp, .x$rate, model_name = 'beta_2012'),
                        upper = get_upper_lims(.x$temp, .x$rate, model_name = 'beta_2012'),
                        supp_errors = 'Y',
                        convergence_count = FALSE)),
         hinshelwood = map(data, ~nls_multstart(rate~hinshelwood_1947(temp = temp, a, e, b, eh),
                        data = .x,
                        iter = c(5,5,5,5),
                        start_lower = get_start_vals(.x$temp, .x$rate, model_name = 'hinshelwood_1947') - 1,
                        start_upper = get_start_vals(.x$temp, .x$rate, model_name = 'hinshelwood_1947') + 1,
                        lower = get_lower_lims(.x$temp, .x$rate, model_name = 'hinshelwood_1947'),
                        upper = get_upper_lims(.x$temp, .x$rate, model_name = 'hinshelwood_1947'),
                        supp_errors = 'Y',
                        convergence_count = FALSE)),
         johnson_lewin = map(data, ~suppressWarnings(nls_multstart(rate~ johnsonlewin_1946(temp = temp, r0, e, eh, topt),
                        data = .x,
                        iter = c(4,4,4,4),
                        start_lower = get_start_vals(.x$temp, .x$rate, model_name = 'johnsonlewin_1946') - 1,
                        start_upper = get_start_vals(.x$temp, .x$rate, model_name = 'johnsonlewin_1946') + 1,
                        lower = get_lower_lims(.x$temp, .x$rate, model_name = 'johnsonlewin_1946'),
                        upper = get_upper_lims(.x$temp, .x$rate, model_name = 'johnsonlewin_1946'),
                        supp_errors = 'Y',
                        convergence_count = FALSE))),
         lactin2 = map(data, ~nls_multstart(rate~lactin2_1995(temp = temp, a, b, tmax, delta_t),
                        data = .x,
                        iter = c(4,4,4,4),
                        start_lower = get_start_vals(.x$temp, .x$rate, model_name = 'lactin2_1995') - 10,
                        start_upper = get_start_vals(.x$temp, .x$rate, model_name = 'lactin2_1995') + 10,
                        lower = get_lower_lims(.x$temp, .x$rate, model_name = 'lactin2_1995'),
                        upper = get_upper_lims(.x$temp, .x$rate, model_name = 'lactin2_1995'),
                        supp_errors = 'Y',
                        convergence_count = FALSE)),
         oneill = map(data, ~nls_multstart(rate~oneill_1972(temp = temp, rmax, ctmax, topt, q10),
                        data = .x,
                        iter = c(4,4,4,4),
                        start_lower = get_start_vals(.x$temp, .x$rate, model_name = 'oneill_1972') - 10,
                        start_upper = get_start_vals(.x$temp, .x$rate, model_name = 'oneill_1972') + 10,
                        lower = get_lower_lims(.x$temp, .x$rate, model_name = 'oneill_1972'),
                        upper = get_upper_lims(.x$temp, .x$rate, model_name = 'oneill_1972'),
                        supp_errors = 'Y',
                        convergence_count = FALSE)),
         pawar = map(data, ~nls_multstart(rate~pawar_2018(temp = temp, r_tref, e, eh, topt, tref = 15),
                        data = .x,
                        iter = c(4,4,4,4),
                        start_lower = get_start_vals(.x$temp, .x$rate, model_name = 'pawar_2018') - 10,
                        start_upper = get_start_vals(.x$temp, .x$rate, model_name = 'pawar_2018') + 10,
                        lower = get_lower_lims(.x$temp, .x$rate, model_name = 'pawar_2018'),
                        upper = get_upper_lims(.x$temp, .x$rate, model_name = 'pawar_2018'),
                        supp_errors = 'Y',
                        convergence_count = FALSE)),
    
         rezende = map(data, ~nls_multstart(rate~rezende_2019(temp = temp, q10, a,b,c),
                        data = .x,
                        iter = c(4,4,4,4),
                        start_lower = get_start_vals(.x$temp, .x$rate, model_name = 'rezende_2019') - 10,
                        start_upper = get_start_vals(.x$temp, .x$rate, model_name = 'rezende_2019') + 10,
                        lower = get_lower_lims(.x$temp, .x$rate, model_name = 'rezende_2019'),
                        upper = get_upper_lims(.x$temp, .x$rate, model_name = 'rezende_2019'),
                        supp_errors = 'Y',
                        convergence_count = FALSE)),
         spain = map(data, ~nls_multstart(rate~spain_1982(temp = temp, a,b,c,r0),
                        data = .x,
                        iter = c(4,4,4,4),
                        start_lower = get_start_vals(.x$temp, .x$rate, model_name = 'spain_1982') - 1,
                        start_upper = get_start_vals(.x$temp, .x$rate, model_name = 'spain_1982') + 1,
                        lower = get_lower_lims(.x$temp, .x$rate, model_name = 'spain_1982'),
                        upper = get_upper_lims(.x$temp, .x$rate, model_name = 'spain_1982'),
                        supp_errors = 'Y',
                        convergence_count = FALSE)),
         thomas2 = map(data, ~nls_multstart(rate~thomas_2017(temp = temp, a,b,c,d,e),
                        data = .x,
                        iter = c(3,3,3,3,3),
                        start_lower = get_start_vals(.x$temp, .x$rate, model_name = 'thomas_2017') - 10,
                        start_upper = get_start_vals(.x$temp, .x$rate, model_name = 'thomas_2017') + 10,
                        lower = get_lower_lims(.x$temp, .x$rate, model_name = 'thomas_2017'),
                        upper = get_upper_lims(.x$temp, .x$rate, model_name = 'thomas_2017'),
                        supp_errors = 'Y',
                        convergence_count = FALSE)))

```


# Overlay multiple TPC models on real data
The predictions of each model can be estimated using broom::augment(). By stacking the models into long format, this can be done on all models at once. To create a smooth curve fit, the predictions are done on a new temperature vector that has 100 points over the temperature range. The predictions for each model formulation are then visualised in ggplot2.

```{r overlay TPCs, echo=FALSE}
# stack models
d_stack <- select(d_fits, -data) %>%
  pivot_longer(., names_to = 'model_name', values_to = 'fit', hinshelwood:thomas2)

# get predictions using augment
newdata <- tibble(temp = seq(min(d$temp), max(d$temp), length.out = 100))
d_preds <- d_stack %>%
  mutate(., preds = map(fit, augment, newdata = newdata)) %>%
  select(-fit) %>%
  unnest(preds)

# take a random point from each model for labelling
d_labs <- filter(d_preds, temp < 30) %>%
  group_by(., model_name) %>%
  sample_n(., 1) %>%
  ungroup()

# plot
ggplot(d_preds, aes(temp, .fitted)) +
  geom_line(aes(col = model_name)) +
  geom_label_repel(aes(temp, .fitted, label = model_name, col = model_name), fill = 'white', nudge_y = 0.8, segment.size = 0.2, segment.colour = 'grey50', d_labs) +
  geom_point(aes(temp, rate), d) +
  theme_bw(base_size = 12) +
  theme(legend.position = 'none') +
  ylim(.45, .65) +
  labs(x = 'Temperature (ºC)',
       y = 'FvFm',
       title = 'Photochemical Efficiency across temperatures') +
  geom_hline(aes(yintercept = 0), linetype = 2) +
  scale_color_brewer(type = 'qual', palette = 2)
```

# Compare Model fits using AICc

1. Several models were omitted that did not fit data at all:
boatman, briere2, joehnk, lrf, ratkowsky, sharpeschoolfull, sharpeschoollow, modified gaussian, kamykowski, gaussian

2. After examining data from several coral genotyples, the most promising TPC models include:
thomas2, spain, rezende, lactin2,pawar, oneill, johnson_lewin, hinshelwood, beta

```{r get AIC, echo = FALSE}
d_ic <- d_stack %>%
  mutate(., info = map(fit, glance),
         AICc =  map_dbl(fit, MuMIn::AICc)) %>%
  select(-fit) %>%
  unnest(info) %>%
  select(model_name, sigma, AIC, AICc, BIC, df.residual)

d_ic
```

# Model selection
Use AICc score to compare between models. For a model selection approach, the model with the lowest AICc score is chosen as the model that best supports the data. The best model varies by genotype but pawar, johnson_lewin, and rezende have shown up as best fitting models according to AICc.

```{r highlight best model, echo= FALSE}
# filter for best model
best_model = filter(d_ic, AICc == min(AICc)) %>% pull(model_name)
best_model 
#> [1] "johnson_lewin" "pawar"  

# get colour code
col_best_mod = RColorBrewer::brewer.pal(n = 6, name = "Dark2")[6]

# plot
ggplot(d_preds, aes(temp, .fitted)) +
  geom_line(aes(group = model_name), col = 'grey50', alpha = 0.5) +
  geom_line(data = filter(d_preds, model_name == best_model), col = col_best_mod) +
  geom_label_repel(aes(temp, .fitted, label = model_name), fill = 'white', nudge_y = 0.8, segment.size = 0.2, segment.colour = 'grey50', data = filter(d_labs, model_name == best_model), col = col_best_mod) +
  geom_point(aes(temp, rate), d) +
  theme_bw(base_size = 12) +
  theme(legend.position = 'none') +
  ylim(.4,.65) +
  labs(x = 'Temperature (ºC)',
       y = 'FvFm',
       title = 'Photochemical efficiency across temperatures')
  geom_hline(aes(yintercept = 0), linetype = 2) 
```

# Model averaging
Not sure if I will average models but here is how it can be done. For a model averaging approach, predictions and parameters from the models are averaged. In ecology, this is usually done based on each model’s weighting. The best supported model’s predictions are taken into account more than the least supported. Often the number of models is reduced by setting a cut-off for the difference in the information criterion metric being used. A common approach is to only keep models within Δ2𝐴𝐼𝐶 of the best model.
```{r, echo = FALSE}
# get model weights
# filtering on AIC score is hashtagged out in this example
d_ic <- d_ic %>%
  # filter(d_ic, aic - min(aic) <= 2) %>%
  mutate(., weight = MuMIn::Weights(AICc))

select(d_ic, model_name, weight) %>%
  arrange(., desc(weight))

# calculate average prediction
ave_preds <- left_join(d_preds, select(d_ic, model_name, weight)) %>%
  group_by(temp) %>%
  summarise(., .fitted = sum(.fitted*weight)) %>%
  ungroup() %>%
  mutate(model_name = 'model average')
#> Joining, by = "model_name"

# create label for averaged predictions
d_labs <- filter(ave_preds, temp < 30) %>% sample_n(., 1)

# plot these
ggplot(d_preds, aes(temp, .fitted)) +
  geom_line(aes(col = model_name), alpha = 0.3) +
  geom_line(data = ave_preds, col = 'blue') +
  geom_label_repel(aes(label = model_name), fill = 'white', nudge_y = 0.8, segment.size = 0.2, segment.colour = 'grey50', data = d_labs, col = 'blue') +
  geom_point(aes(temp, rate), d) +
  theme_bw(base_size = 12) +
  theme(legend.position = 'none') +
  ylim(.40,.65 ) +
  labs(x = 'Temperature (ºC)',
       y = 'FvFm',
       title = 'Respiration across temperatures',
       subtitle= 'Model averaged predictions') +
  geom_hline(aes(yintercept = 0), linetype = 2) +
  scale_color_brewer(type = 'qual', palette = 2)
```


# Get Parameter Estimates and make observations
Cheryl's Observations aftering running through several individual curves:
1. Many of the models do not give a ctmin (Inf) or thermal tolerance, but can still calculate thermal breadth
2. 

```{r get params, echo=FALSE}
# calculate estimated parameters
params <- d_stack %>%
  mutate(., params = map(fit, calc_params)) %>%
  select(-fit) %>%
  unnest(params)

# get averaged parameters based on model weights
ave_params <- left_join(params, select(d_ic, model_name, weight)) %>%
  summarise(., across(rmax:skewness, function(x){sum(x*.$weight)})) %>%
  mutate(model_name = 'model average')
#> Joining, by = "model_name"

# show them
bind_rows(select(params, model_name, rmax:skewness), ave_params) %>%
  mutate_if(is.numeric, round, 2)
```
  
  
```{r view data}
#This gives us a dataframe with our grouping variables
glimpse(select(d_fits, 1:7))
#Each column containing the model stores the actual model fit.
d_fits$beta[[1]]
```

# Plot indiviudal TPCs and get TPC parameters
The parameters from each model fit can be extracted using broom::tidy(). However, as each model has parameters with different meanings, this may not be very useful in this instance.

The predictions of each model can be estimated using broom::augment(). This can be done on all models at once after the models are stacked into long format. To create a smooth curve fit, the predictions are done on a new temperature vector that has 100 points over the temperature range. The predictions for each model formulation are then visualised in ggplot2.

```{r get parameters, echo = FALSE}
# stack models
d_stack <- select(d_fits, -data) %>%
  pivot_longer(., names_to = 'model_name', values_to = 'fit', hinshelwood:thomas2)

# get parameters using tidy
params <- d_stack %>%
  mutate(., est = map(fit, tidy)) %>%
  select(-fit) %>%
  unnest(est)
glimpse(params)


# get predictions using augment
newdata <- tibble(temp = seq(min(d$temp), max(d$temp), length.out = 100))
d_preds <- d_stack %>%
  mutate(., preds = map(fit, augment, newdata = newdata)) %>%
  select(-fit) %>%
  unnest(preds)


```
 
# Plot many TPC models to same data
```{r plot TPCs, echo = FALSE}
# plot
ggplot(d_preds, aes(temp, rate)) +
  geom_point(aes(temp, rate), d) +
  geom_line(aes(temp, .fitted), col = 'blue') +
  facet_wrap(~model_name, labeller = labeller(model_name = label_facets_num), scales = 'free', ncol = 5) +
  theme_bw(base_size = 12) +
  theme(legend.position = 'none',
        strip.text = element_text(hjust = 0),
        strip.background = element_blank()) +
  labs(x = 'Temperature (ºC)',
       y = 'FvFm',
       title = 'Fits of better models available in rTPC') +
  ylim(.45, .65) +
  geom_hline(aes(yintercept = 0), linetype = 2)

# save
filename=paste0('rTPC_fits_Isabela_better_', today.date, '.','png')
ggsave(path= "/Users/loga8761/github/PAM/Figures", filename, width = 8, height = 4)
```
