---
title: "rTPC_v2"
author: "Cheryl Logan"
date: '2022-06-14'
output: html_document
---

```{r setup, include=FALSE}
# load packages
remotes::install_github("padpadpadpad/rTPC")
library(rTPC)
library(nls.multstart)
library(broom)
library(tidyverse)
```

Following 
https://padpadpadpad.github.io/rTPC/articles/rTPC.html

```{r cars}
get_model_names()
# choose model
mod = 'spain_1982'
```

Load PAM data
```{r}
# Load data 

# from "Galapagos_PAM_allsites.Rmd"
# Remove sites with odd data or different temp profiles
PAM_all <- read_csv("/Users/loga8761/github/PAM/Data/Galapagos_all.csv") %>%
  filter(Site != "AcademyBay"& Site != "Floreana") %>%
  mutate(Site = fct_relevel(Site, 
            "Darwin", "Wolf2", "Punta_Pitt", 
            "Espanola", "Isabela"))  

pam.data <- PAM_all %>%
  filter(Species == params$Species) %>%
  #filter(RampType == params$RampType) %>%                               # Replacing values
  mutate_at("Site",str_replace,"Wolf2", "Wolf") %>%                               # Replacing values
  mutate_at("Site",str_replace,"Punta_Pitt", "San Cristobal") %>%  
  select(Geno,Treatment, PAM, Site)

pam.data$Geno <- as.numeric(pam.data$Geno)
pam.data$Site <-  as.factor(pam.data$Site)
pam.data$Treatment <- as.numeric(pam.data$Treatment)
  
#Labels for re-ordering site names (warm -> cold)
sitelabs <- c("Darwin", "Wolf2", "Punta_Pitt", "Espanola", "Isabela")
names(sitelabs) <- c("Darwin", "Wolf","San Cristobal", "Espanola", "Isabela")

# Get subset of data
Darwin<-pam.data %>% 
  filter(Site == "Darwin")  %>%
  arrange(Geno)

#rename columns to follow tutorial more easily
Darwin <- as.data.frame(Darwin)  %>%
  rename(temp = Treatment)  %>%
  rename(rate = PAM)  
```

```{r pressure, echo=FALSE}
# load in data
data("chlorella_tpc")

# keep just a single curve
d <- filter(chlorella_tpc, curve_id == 1)
d <- filter(Darwin, Geno == 2)


# show the data
ggplot(d, aes(temp, rate)) +
  geom_point() +
  theme_bw(base_size = 12) +
  labs(x = 'Temperature (ºC)',
       y = 'FvFm',
       title = 'Photosynthetic efficiency across temperature')
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
# get start values and fit model
start_vals <- get_start_vals(d$temp, d$rate, model_name = 'spain_1982')

# fit model
mod <- nls.multstart::nls_multstart(rate~spain_1982(temp = temp, a, b, c, r0),
data = d,
iter = c(3,3,3,3),
start_lower = start_vals - 1,
start_upper = start_vals + 1,
lower = get_lower_lims(d$temp, d$rate, model_name = 'spain_1982'),
upper = get_upper_lims(d$temp, d$rate, model_name = 'spain_1982'),
supp_errors = 'Y',
convergence_count = FALSE)
```

Using nls_multstart, we will use a random-search/shotgun approach to fit the curve. Random start parameter values are picked from a uniform distribution between start_lower and start_upper for each parameter. If the best model is not improved upon (in terms of AIC score) for 100 new start parameter combinations, the function will return that model fit. This is controlled by convergence_count, if this is set to FALSE, nls_multstart() will try and fit all iterations.


```{r}
# look at model fit
summary(mod)
```

```{r}
# get predictions
preds <- data.frame(temp = seq(min(d$temp), max(d$temp), length.out = 100))
preds <- broom::augment(mod, newdata = preds)

# plot
ggplot(preds) +
geom_point(aes(temp, rate), d) +
geom_line(aes(temp, .fitted), col = 'blue') +
theme_bw()+
  labs(x = 'Temperature (ºC)',
       y = 'FvFm',
       title = 'Photosynthetic efficiency across temperature')
```
```{r}
# get predictions
preds <- data.frame(temp = seq(min(d$temp), max(d$temp), length.out = 100))
preds <- broom::augment(mod, newdata = preds)

# plot
ggplot(preds) +
geom_point(aes(temp, rate), d) +
geom_line(aes(temp, .fitted), col = 'blue') +
theme_bw()+
labs(x = 'Temperature (ºC)',
       y = 'FvFm',
       title = 'Photosynthetic efficiency across temperature')
```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
